---
title: "blh_synthetic"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{blh_synthetic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=6
)
```

# # BLH synthetic example

First we load the package.

```{r setup}
library(HydroTomo)
```

First we want to know that the mean value of the oHT (but notice it could be smaller than usual since it may not stable).

```{r calculate the mean of oHT}
data("oHT")
dfoHT = do.call(rbind,oHT)
mean_oHT = mean(dfoHT$data)
print(paste("The mean value of oHT is ", round(mean_oHT,4)))
```

Then we set the random seed for reproducibility and generate a true transmissivity field.

```{r generate_trueK}
set.seed(200)
data('qHT')
domain=c(80,80,0,80,0,80)
grid <- GenGrid(domain=domain)
# exp(3.5)=33.12 m2/d means very large T values in sand and gravel aquifer.
geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0)
trueK <- random2d(nsim=1,grid=grid,geo = geo)
plot1 <- Plotparameter2d(trueK,iflog=TRUE)
print(plot1)
```

# Generating synthetic data

We then simulate steady-state drawdown data for a pumping test using the true transmissivity field.

```{r generate_synthetic_data}
TT <- trueK[,-c(1,2)]
qHT1 = qHT[c(1,3,5,7,9,11)]
oHT1 = oHT[c(1,3,5,7,9,11)]
### loop over the list to update the value in oHT1.

for (i in 1:length(oHT1)){
  Qinf <- qHT1[[i]]
  trueh <- Fsteady2dsim(TT=TT,Qinf=Qinf,grid=grid,lrw=320000)$solution
  oinf <- oHT1[[i]]
  oinf = samData(grid = grid,Oinf = oinf,h=trueh)
  oHT1[[i]] <- oinf[,c('data','x','y')]
}

dfoHT1 = do.call(rbind,oHT1)
mean_oHT1 = mean(dfoHT1$data)
print(paste("The mean value of oHT1 is ", round(mean_oHT1,4)))
```


Now we can save the data in the package usethis:: use_data. We create the object list that save trueK, qHT1, oHT1, and grid.

```{r save_data, eval=FALSE}
blh_synthetic <- list(trueK=trueK, qHT=qHT1, oHT=oHT1, grid=grid)
usethis::use_data(blh_synthetic, overwrite = TRUE)
```
Now the data is saved in the package and can be loaded using data('blh_synthetic').

# Performing inversion

We can then perform inversion using the synthetic data.

```{r perform_inversion, cache=TRUE}
result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT1, nsim=50, itermax=10, lrw=320000)
```

# Plotting inversion results

```{r plot_inversion_results}
inversePlot(niterm=7,grid=grid,iterdf = result,oHT = oHT1,trueK=trueK)
```

# noised data

The head match perfectly, but in reality the head data often contains noise. We can add noise to the synthetic data and perform inversion again.

```{r add_noise_and_invert, cache=TRUE}
set.seed(300)
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}
result_noised <- Finverse(grid =grid, qHT = qHT1, oHT = oHT_noised, nsim=50, itermax=10, lrw=320000)
inversePlot(niterm=7,grid=grid,iterdf = result_noised,oHT = oHT_noised,trueK=trueK)
```

Now we increase the sd of the noise to see how it affects the inversion results.

```{r increase_noise_and_invert, cache=TRUE}
set.seed(400)
oHT_noised2 <- oHT1
for (i in 1:length(oHT_noised2)){
  noise = rnorm(n=nrow(oHT_noised2[[i]]), mean=0, sd=0.2*sd(oHT_noised2[[i]]$data))
  oHT_noised2[[i]]$data <- oHT_noised2[[i]]$data + noise
}
result_noised2 <- Finverse(grid =grid, qHT = qHT1, oHT = oHT_noised2, nsim=50, itermax=10, lrw=320000)
inversePlot(niterm=10,grid=grid,iterdf = result_noised2,oHT = oHT_noised2,trueK=trueK)
```

This gives an idea of how noise in the data affects the inversion results. Now we look at if reducing the number of pumping tests affects the inversion results.

```{r reduce_tests_and_invert, cache=TRUE}
qHT_reduced = qHT1[c(1,5)]
oHT_reduced = oHT1[c(1,5)]
result_reduced <- Finverse(grid =grid, qHT = qHT_reduced, oHT = oHT_reduced, nsim=50, itermax=10, lrw=320000)
inversePlot(niterm=10,grid=grid,iterdf = result_reduced,oHT = oHT_reduced,trueK=trueK)
```
