---
title: "Sensitivity and Cross-covariance: Single Well Pair"
author:
  - name: "Yuanyuan Zha"
    affiliation: "Wuhan University"
    email: "zhayuan87@whu.edu.cn"
    orcid: "0000-0003-4323-0730"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    fig_width: 10
    fig_height: 7
subtitle: "Adjoint vs Monte Carlo — Sensitivity & Cross-covariance"
vignette: >
  %\VignetteIndexEntry{Sensitivity and Cross-covariance: Single Well Pair}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo       = TRUE,
  warning    = FALSE,
  message    = FALSE,
  collapse   = TRUE,
  comment    = "#>",
  fig.width  = 10,
  fig.height = 7,
  cache      = FALSE
)
library(ggplot2)
library(patchwork)
```

# 1. Problem Setup

40 × 40 grid, pumping well at (25, 50), observation well at (75, 50).

```{r synthetic-setup}
library(HydroTomo)
set.seed(42)

# ---- Grid ----
domain <- c(40, 40, 0, 100, 0, 100)
grid   <- GenGrid(domain)
nx     <- grid$nx
ny     <- grid$ny
nelem  <- nx * ny
cat(sprintf("Grid: %d × %d,  %d elements\n", nx, ny, nelem))

# ---- True transmissivity field ----
trueK <- random2d(grid = grid, nsim = 1,
                   geo = list(me = 0, var = 1, geomod = "Gau",
                              anis = c(90, 1), range = 20, nugget = 0.01))
TT_true <- trueK$Tp

# ---- Pre-built grid data frame for ggplot ----
# image() convention: matrix row = y, column = x
# ggplot geom_raster: x = column index, y = row index
df_map <- expand.grid(col = 1:ny, row = 1:nx)

vec2df <- function(v, df = df_map, name = "z") {
  m <- matrix(v, nrow = ny, ncol = nx)
  df[[name]] <- m[cbind(df$col, df$row)]
  df
}

# Plot true ln(T) field
p_lnT <- ggplot(vec2df(log(trueK$Tp)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_viridis_c(name = "ln(T)") +
  coord_fixed(expand = FALSE) +
  labs(title = "True ln(T) field", x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(panel.grid = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))
p_lnT
```

```{r wells-setup}
# ---- Pumping test ----
Qinf1 <- data.frame(Qp = 10, x = 25.5, y = 50.5)
qHT <- list(test1 = Qinf1)

# Forward simulation to obtain "true" head
trueh <- Fsteady2dsim(grid = grid, TT = TT_true, Qinf = Qinf1)

# ---- Observation well ----
Oinf1 <- data.frame(data = NA, x = 75.5, y = 50.5)
Oinf1 <- samData(grid = grid, Oinf = Oinf1, h = trueh$solution)
oHT   <- list(test1 = Oinf1)
nobs  <- nrow(Oinf1)

cat(sprintf("Pumping well: (%.1f, %.1f)\n", Qinf1$x, Qinf1$y))
cat(sprintf("Observation  : (%.1f, %.1f)\n", Oinf1$x[1], Oinf1$y[1]))
cat(sprintf("Observed drawdown: %.4f\n", Oinf1$data[1]))

# ---- Well indices (image coordinates) ----
pcol <- Qinf1$x / domain[4] * ny   # pumping well column index
prow <- Qinf1$y / domain[6] * nx   # pumping well row index
ocol <- Oinf1$x[1] / domain[4] * ny # observation well column index
orow <- Oinf1$y[1] / domain[6] * nx # observation well row index

# Plot drawdown field + well locations
df_dd <- vec2df(trueh$solution)
p_dd <- ggplot(df_dd, aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_viridis_c(name = "Drawdown") +
  geom_point(aes(x = pcol, y = prow, shape = "Pumping well"),
             color = "red", size = 3) +
  geom_point(aes(x = ocol, y = orow, shape = "Observation well"),
             color = "blue", size = 3) +
  scale_shape_manual(name = NULL,
    values = c("Pumping well" = 17, "Observation well" = 4)) +
  coord_fixed(expand = FALSE) +
  labs(title = "Drawdown field", x = NULL, y = NULL) +
  theme_minimal(base_size = 12) +
  theme(panel.grid = element_blank(),
        axis.text  = element_blank(),
        axis.ticks = element_blank(),
        plot.title = element_text(hjust = 0.5))
p_dd
```

# 2. Adjoint: Sensitivity $\mathbf{J}$ and Cross-covariance $\mathbf{C}_{hk}$

```{r adjoint, cache=TRUE}
geo0 <- list(me = 0, var = 1, geomod = "Gau",
             anis = c(90, 1), range = 20, nugget = 0)

# Adjoint Jacobian matrix
J_adj <- jacobian2D(grid = grid, TT = exp(geo0$me + geo0$var/2),
                    qHT = qHT, oHT = oHT)
  
  


# J_adj: nobs × nelem (nobs = 1 here)

# Build prior covariance C_kk
xy      <- grid$grid
dist_mat <- as.matrix(dist(xy))
C_kk <- geo0$var * exp(-dist_mat / geo0$range)

# Adjoint cross-covariance
C_hk_adj <- FinverseAdj(
  grid  = grid, qHT = qHT, oHT = oHT,
  ifcor = TRUE, geo  = geo0
)

cat(sprintf("J_adj dimension:    %d × %d\n", nrow(J_adj), ncol(J_adj)))
cat(sprintf("C_hk_adj dimension: %d × %d\n", nrow(C_hk_adj), ncol(C_hk_adj)))
```

# 3. Monte Carlo: Direct $\mathbf{C}_{hk}$ estimation

```{r mc-covhk, cache=TRUE}
# Normal correlation length → direct cross-covariance
mc_cov <- Finverse(
  grid  = grid, qHT = qHT, oHT = oHT,
  nsim  = 200, ifcor = TRUE, geo  = geo0
)
# mc_cov: C_hk, nobs × nelem
dim(mc_cov)
```

# 4. Monte Carlo approximation of sensitivity

```{r mc-sensitivity, cache=TRUE}
# Tiny correlation length: range → 0 → C_kk ≈ σ²I → C_hk ≈ σ²J
geo_tiny <- list(me = 0, var = 1, geomod = "Gau",
                 anis = c(90, 1), range = 1, nugget = 0)

set.seed(2024)
mc_tiny <- Finverse(
  grid  = grid, qHT = qHT, oHT = oHT,
  nsim  = 200, ifcor = TRUE, geo  = geo_tiny
)

# C_hk ≈ var · J
J_mc <- mc_tiny / geo_tiny$var
dim(J_mc)
```

---

# 5. Comparison: Sensitivity $\mathbf{J}$

## 5.1 Side-by-side comparison

```{r sens-comparison, fig.width=12, fig.height=5.5}
gmax_sens <- max(max(abs(J_adj)), max(abs(J_mc)))

# Shared well points layer
geom_wells <- list(
  geom_point(aes(x = pcol, y = prow, shape = "Pumping"),
             color = "green3", size = 2.5),
  geom_point(aes(x = ocol, y = orow, shape = "Obs."),
             color = "red",    size = 2.5),
  scale_shape_manual(name = NULL,
    values = c("Pumping" = 17, "Obs." = 4))
)

theme_map <- theme_minimal(base_size = 11) +
  theme(panel.grid   = element_blank(),
        axis.text    = element_blank(),
        axis.ticks   = element_blank(),
        plot.title   = element_text(hjust = 0.5, size = 12),
        legend.position = "right")

p1 <- ggplot(vec2df(as.vector(J_adj)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = "J",
                       limits = c(-gmax_sens, gmax_sens)) +
  geom_wells +
  coord_fixed(expand = FALSE) +
  labs(title = "Adjoint J (sensitivity)", x = NULL, y = NULL) +
  theme_map

p2 <- ggplot(vec2df(as.vector(J_mc)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = "J",
                       limits = c(-gmax_sens, gmax_sens)) +
  geom_wells +
  coord_fixed(expand = FALSE) +
  labs(title = expression(MC~J~(r %->% 0*","~N == 200)), x = NULL, y = NULL) +
  theme_map

p1 + p2 + plot_annotation(tag_levels = "a")
```

## 5.2 Profile along x

```{r sens-profile-x, fig.width=12, fig.height=5}
# Extract profile at y ≈ 50
y_idx <- round(orow)

# (a) 2D sensitivity map + profile line
p2d <- ggplot(vec2df(as.vector(J_adj)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = "J",
                       limits = c(-gmax_sens, gmax_sens)) +
  geom_hline(yintercept = y_idx, color = "black", linewidth = 1, linetype = "dashed") +
  geom_point(aes(x = pcol, y = prow), shape = 17, color = "green3", size = 2.5) +
  geom_point(aes(x = ocol, y = orow), shape = 4,  color = "red",    size = 2.5) +
  coord_fixed(expand = FALSE) +
  labs(title = "Adjoint J (y ≈ 50 profile)", x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid   = element_blank(),
        axis.text    = element_blank(),
        axis.ticks   = element_blank(),
        plot.title   = element_text(hjust = 0.5, size = 12),
        legend.position = "right")

# (b) Profile — reshape 1×nelem row vectors before extraction
J_adj_mat <- matrix(as.vector(J_adj), nrow = ny, ncol = nx)
J_mc_mat  <- matrix(as.vector(J_mc),  nrow = ny, ncol = nx)
x_seq <- (1:ny) / ny * domain[4]
df_prof <- data.frame(
  x    = x_seq,
  adj  = J_adj_mat[, y_idx],
  mc   = J_mc_mat[, y_idx]
)

p_prof <- ggplot(df_prof, aes(x)) +
  geom_line(aes(y = adj, color = "Adjoint J"),   linewidth = 1) +
  geom_line(aes(y = mc,  color = "MC J (r→0)"), linewidth = 1, linetype = "dashed") +
  geom_vline(xintercept = c(25, 75), linetype = "dotted", color = "gray50") +
  scale_color_manual(name = NULL,
    values = c("Adjoint J" = "steelblue", "MC J (r→0)" = "tomato")) +
  labs(x = "x", y = "Sensitivity J",
       title = "J along y ≈ 50") +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5))

p2d + p_prof + plot_annotation(tag_levels = "a")
```

## 5.3 Scatter: cell-by-cell comparison

```{r sens-scatter, fig.width=7, fig.height=6}
cor_sens <- cor(as.vector(J_adj), as.vector(J_mc))
df_scatter <- data.frame(adj = as.vector(J_adj), mc = as.vector(J_mc))

ggplot(df_scatter, aes(adj, mc)) +
  geom_point(alpha = 0.15, size = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red",
              linewidth = 1, linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE, color = "blue", linewidth = 1.2) +
  annotate("text", x = min(df_scatter$adj), y = max(df_scatter$mc),
           label = sprintf("r = %.4f", cor_sens),
           hjust = 0, vjust = 1, size = 5) +
  labs(title = "Sensitivity: adjoint vs MC",
       x = "Adjoint J", y = "MC J (r → 0)") +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(hjust = 0.5))
```

---

# 6. Comparison: Cross-covariance $\mathbf{C}_{hk}$

## 6.1 Side-by-side comparison

```{r covhk-comparison, fig.width=12, fig.height=5.5}
gmax_cov <- max(max(abs(C_hk_adj)), max(abs(mc_cov)))

p1c <- ggplot(vec2df(as.vector(C_hk_adj)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = expression(C[hk]),
                       limits = c(-gmax_cov, gmax_cov)) +
  geom_wells +
  coord_fixed(expand = FALSE) +
  labs(title = expression(Adjoint~C[hk]~(covariance)), x = NULL, y = NULL) +
  theme_map

p2c <- ggplot(vec2df(as.vector(mc_cov)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = expression(C[hk]),
                       limits = c(-gmax_cov, gmax_cov)) +
  geom_wells +
  coord_fixed(expand = FALSE) +
  labs(title = expression(MC~C[hk]~(N == 200)), x = NULL, y = NULL) +
  theme_map

p1c + p2c + plot_annotation(tag_levels = "a")
```

## 6.2 Profile along x

```{r covhk-profile-x, fig.width=12, fig.height=5}
# (a) 2D covariance map + profile line
p2d_c <- ggplot(vec2df(as.vector(C_hk_adj)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = expression(C[hk]),
                       limits = c(-gmax_cov, gmax_cov)) +
  geom_hline(yintercept = y_idx, color = "black", linewidth = 1, linetype = "dashed") +
  geom_point(aes(x = pcol, y = prow), shape = 17, color = "green3", size = 2.5) +
  geom_point(aes(x = ocol, y = orow), shape = 4,  color = "red",    size = 2.5) +
  coord_fixed(expand = FALSE) +
  labs(title = expression(Adjoint~C[hk]~(y %~~% 50~profile)),
       x = NULL, y = NULL) +
  theme_minimal(base_size = 11) +
  theme(panel.grid   = element_blank(),
        axis.text    = element_blank(),
        axis.ticks   = element_blank(),
        plot.title   = element_text(hjust = 0.5, size = 12),
        legend.position = "right")

# (b) Profile — reshape 1×nelem row vectors before extraction
C_hk_adj_mat <- matrix(as.vector(C_hk_adj), nrow = ny, ncol = nx)
mc_cov_mat   <- matrix(as.vector(mc_cov),   nrow = ny, ncol = nx)
df_prof_c <- data.frame(
  x   = x_seq,
  adj = C_hk_adj_mat[, y_idx],
  mc  = mc_cov_mat[, y_idx]
)

p_prof_c <- ggplot(df_prof_c, aes(x)) +
  geom_line(aes(y = adj, color = "Adjoint C_hk"), linewidth = 1) +
  geom_line(aes(y = mc,  color = "MC C_hk"),      linewidth = 1, linetype = "dashed") +
  geom_vline(xintercept = c(25, 75), linetype = "dotted", color = "gray50") +
  scale_color_manual(name = NULL,
    values = c("Adjoint C_hk" = "steelblue", "MC C_hk" = "tomato")) +
  labs(x = "x", y = expression(C[hk]),
       title = expression(C[hk]~along~y %~~% 50)) +
  theme_minimal(base_size = 11) +
  theme(plot.title = element_text(hjust = 0.5))

p2d_c + p_prof_c + plot_annotation(tag_levels = "a")
```

## 6.3 Scatter: cell-by-cell comparison

```{r covhk-scatter, fig.width=7, fig.height=6}
cor_cov <- cor(as.vector(C_hk_adj), as.vector(mc_cov))
df_scatter_c <- data.frame(adj = as.vector(C_hk_adj), mc = as.vector(mc_cov))

ggplot(df_scatter_c, aes(adj, mc)) +
  geom_point(alpha = 0.15, size = 0.5) +
  geom_abline(slope = 1, intercept = 0, color = "red",
              linewidth = 1, linetype = "dashed") +
  geom_smooth(method = "lm", se = FALSE, color = "blue", linewidth = 1.2) +
  annotate("text", x = min(df_scatter_c$adj), y = max(df_scatter_c$mc),
           label = sprintf("r = %.4f", cor_cov),
           hjust = 0, vjust = 1, size = 5) +
  labs(title = "Cross-covariance: adjoint vs MC",
       x = expression(Adjoint~C[hk]), y = expression(MC~C[hk])) +
  theme_minimal(base_size = 12) +
  theme(plot.title = element_text(hjust = 0.5))
```

---

# 7. Sensitivity vs Covariance: shape comparison

Smoothing effect of $\mathbf{C}_{kk}$: sensitivity maps (sharp) vs covariance maps (smooth) side by side.

```{r sens-vs-covhk, fig.width=16, fig.height=4.5}
# Independent scales for sensitivity and covariance (legends may differ)
theme_4panel <- theme_minimal(base_size = 10) +
  theme(panel.grid    = element_blank(),
        axis.text     = element_blank(),
        axis.ticks    = element_blank(),
        plot.title    = element_text(hjust = 0.5, size = 10),
        legend.position = "right",
        legend.key.height = unit(0.7, "cm"))

pA <- ggplot(vec2df(as.vector(J_adj)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = "J",
                       limits = c(-gmax_sens, gmax_sens)) +
  geom_point(aes(x = pcol, y = prow), shape = 17, color = "green3", size = 1.5) +
  geom_point(aes(x = ocol, y = orow), shape = 4,  color = "red",    size = 1.5) +
  coord_fixed(expand = FALSE) +
  labs(title = "Adjoint J\n(sensitivity, sharp)", x = NULL, y = NULL) +
  theme_4panel

pB <- ggplot(vec2df(as.vector(J_mc)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = "J",
                       limits = c(-gmax_sens, gmax_sens)) +
  geom_point(aes(x = pcol, y = prow), shape = 17, color = "green3", size = 1.5) +
  geom_point(aes(x = ocol, y = orow), shape = 4,  color = "red",    size = 1.5) +
  coord_fixed(expand = FALSE) +
  labs(title = "MC J (r → 0)\n(sensitivity, sharp)", x = NULL, y = NULL) +
  theme_4panel

pC <- ggplot(vec2df(as.vector(C_hk_adj)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = expression(C[hk]),
                       limits = c(-gmax_cov, gmax_cov)) +
  geom_point(aes(x = pcol, y = prow), shape = 17, color = "green3", size = 1.5) +
  geom_point(aes(x = ocol, y = orow), shape = 4,  color = "red",    size = 1.5) +
  coord_fixed(expand = FALSE) +
  labs(title = expression(Adjoint~C[hk]~(covariance~smooth)),
       x = NULL, y = NULL) +
  theme_4panel

pD <- ggplot(vec2df(as.vector(mc_cov)), aes(col, row, fill = z)) +
  geom_raster() +
  scale_fill_distiller(palette = "RdBu", name = expression(C[hk]),
                       limits = c(-gmax_cov, gmax_cov)) +
  geom_point(aes(x = pcol, y = prow), shape = 17, color = "green3", size = 1.5) +
  geom_point(aes(x = ocol, y = orow), shape = 4,  color = "red",    size = 1.5) +
  coord_fixed(expand = FALSE) +
  labs(title = expression(MC~C[hk]~(N == 200)~(covariance~smooth)),
       x = NULL, y = NULL) +
  theme_4panel

(pA | pB | pC | pD) + plot_annotation(tag_levels = "a")
```

---

# 8. Summary

```{r summary}
cat("========================================\n")
cat("  Sensitivity (J) comparison\n")
cat(sprintf("  Pearson r = %.4f\n", cor_sens))
cat("========================================\n")
cat("\n")
cat("========================================\n")
cat("  Cross-covariance (C_hk) comparison\n")
cat(sprintf("  Pearson r = %.4f\n", cor_cov))
cat("========================================\n")
```

| Metric | Adjoint | Monte Carlo (N = 200) |
|------|:-----------:|:---------------------:|
| $\mathbf{J}$ (sensitivity) shape | Sharp, local (concentrated near well-pair line) | Sharp, local (approaches adjoint) |
| $\mathbf{C}_{hk}$ (covariance) shape | Smooth, spread out (smoothed by range = 20) | Smooth, spread out |
| $\mathbf{J}$ correlation coefficient | — | `r round(cor_sens, 4)` |
| $\mathbf{C}_{hk}$ correlation coefficient | — | `r round(cor_cov, 4)` |

- **Sensitivity** reflects the direct impact of a single element's parameter on the observed head, forming high-gradient zones near the pumping-observation well pair line.
- **Cross-covariance** spreads the sensitivity spatially through the prior covariance $\mathbf{C}_{kk}$, producing more physically plausible update weights.
- The Monte Carlo and adjoint methods agree closely at both the sensitivity and covariance levels, validating the relationship $\mathbf{C}_{hk} = \mathbf{J} \mathbf{C}_{kk}$.
