---
title: "ht9_synthetic"
author:
  - name: "Yuanyuan Zha"
    affiliation: "Wuhan University"
    email: "zhayuan87@whu.edu.cn"
    orcid: "0000-0003-4323-0730"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ht9_synthetic}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=6,
  fig.height=6
)
```

# HT 9-well synthetic example

This vignette demonstrates a simple 9-well hydraulic tomography setup on a 9 * 9 grid.
Nine wells are placed at grid positions 3, 5, 7 in both x and y directions (3 * 3 layout).
All 9 sequential pumping tests are performed: when well *i* pumps, the remaining 8 wells observe.

First we load the package.

```{r setup}
library(HydroTomo)
```

## Generate true transmissivity field

We set the random seed for reproducibility and generate a log-normal transmissivity
field on a 9&times;9 grid.

```{r generate_trueK}
set.seed(100)
domain <- c(9, 9, 0, 9, 0, 9)
grid <- GenGrid(domain = domain)

# Moderately heterogeneous aquifer: mean ln(T) = 3 (T ~ 20 m2/d)
geo <- list(me = 3, var = 1, geomod = "Exp", anis = c(90, 1), range = 6, nugget = 0)
trueK <- random2d(nsim = 1, grid = grid, geo = geo)
Plotparameter2d(trueK, iflog = TRUE)
```

## Well layout

Wells are placed at grid-cell centres 3, 5, 7 in both x and y, giving
9 wells on a regular 3&times;3 grid.

```{r well_layout}
# Grid-cell centres: (cell_num - 0.5) * cell_width
well_coords <- expand.grid(x = c(2.5, 4.5, 6.5), y = c(2.5, 4.5, 6.5))
well_names <- paste0("W", 1:9)
well_coords$well <- well_names
print(well_coords)
```

## Build pumping and observation lists

For each of the 9 wells, we create one pumping-test entry and one observation
entry containing all other 8 wells.

```{r build_qHT_oHT}
TT <- trueK[, -c(1, 2)]

qHT <- list()
oHT <- list()

for (i in 1:9) {
  test_name <- paste0("test_W", i)

  # Pumping well i
  Qinf <- data.frame(
    Qp = 10,               # pumping rate [L3/T]
    xp = well_coords$x[i],
    yp = well_coords$y[i]
  )

  # Observation: all wells except the pumping one
  obs_wells <- well_coords[-i, ]
  Oinf <- data.frame(
    data = NA,
    x    = obs_wells$x,
    y    = obs_wells$y
  )

  qHT[[test_name]] <- Qinf
  oHT[[test_name]] <- Oinf
}
```

## Simulate steady-state drawdown

We run a steady-state forward simulation for each pumping test using the true
transmissivity field, then sample the drawdown at observation locations.

```{r simulate_drawdown}
for (i in 1:9) {
  Qinf <- qHT[[i]]
  trueh <- Fsteady2dsim(TT = TT, Qinf = Qinf, grid = grid)$solution
  oinf <- oHT[[i]]
  oinf <- samData(grid = grid, Oinf = oinf, h = trueh)
  oHT[[i]] <- oinf[, c("data", "x", "y")]
}

# Quick summary of synthetic drawdowns
dfoHT <- do.call(rbind, oHT)
print(paste("Drawdown range:", round(min(-dfoHT$data), 4),
            "to", round(max(-dfoHT$data), 4),
            "  mean =", round(mean(-dfoHT$data), 4)))
```

## Package data object

The `HTdata` list bundles the true field, pumping configuration, observations,
and grid into a single object that can be stored in the package.

```{r save_data, eval=FALSE}
ht9_synthetic <- list(trueK = trueK, qHT = qHT, oHT = oHT, grid = grid)
# usethis::use_data(ht9_synthetic, overwrite = TRUE)
```

## Inversion (no noise)

We perform ensemble-smoother-based Bayesian inversion using first 5 pumping tests.
With 9 tests on a small 9 * 9 grid, convergence is rapid.

```{r inversion_clean, cache=TRUE}
n = 9
result <- Finverse(
  grid   = grid,
  qHT    = qHT[1:n],
  oHT    = oHT[1:n],
  nsim   = 200,
  itermax = 10,
  lrw    = 160000,
  decay  = 2
)
```

```{r plot_inversion_clean}
inversePlot(niterm = 10, grid = grid, iterdf = result, oHT = oHT[1:n], trueK = trueK)
```
