---
title: "HydroTomo - Hydraulic Tomography Analysis (Adjoint Method)"
author: ''
date: "`r Sys.Date()`"
output:
  word_document:
    toc: true
  html_document:
    toc: true
    toc_float: true
    code_folding: show
    fig_width: 10
    fig_height: 7
subtitle: "Steady-State 3D Adjoint-Based Inversion with Well Screens"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE,
                      fig.width = 10, fig.height = 7)
```

# 1. Load Packages

```{r packages}
library(HydroTomo)
library(ggplot2)
library(dplyr)
library(tidyr)
library(gridExtra)
```

# 2. Read and Inspect CSV Data

```{r read_csv}
# File paths
data_dir <- file.path("D:", "BaiduSyncdisk", "研究生", "董海洲",
                       "水力层析抽水试验数据 (7.24)", "水力层析抽水试验数据")

# Read all CSVs
wells <- read.csv(file.path(data_dir, "wells_xyz.csv"),
                  stringsAsFactors = FALSE)
drawdown <- read.csv(file.path(data_dir, "Final_Drawdown_Summary.csv"),
                     stringsAsFactors = FALSE, fileEncoding = "UTF-8")
pump_summary <- read.csv(file.path(data_dir, "pumping_test_summary.csv"),
                         stringsAsFactors = FALSE, fileEncoding = "UTF-8")
pump_rate <- read.csv(file.path(data_dir, "pumping_rate_info.csv"),
                      stringsAsFactors = FALSE, fileEncoding = "UTF-8")

# Inspect
cat("=== wells_xyz ===\n")
print(wells)

cat("\n=== pumping_rate_info ===\n")
print(pump_rate)

cat("\n=== pumping_test_summary ===\n")
print(pump_summary)

cat("\n=== Final_Drawdown_Summary (unique tests) ===\n")
print(unique(drawdown[[1]]))
```

# 3. Coordinate Translation and Well Classification

HydroTomo convention: z downward positive (depth).  
Exclude distant wells K1-K4, focus on SY1-SY5 cluster for observation.

```{r coord_translate}
# ---- Well classification ----
wells$type <- ifelse(grepl("^SY", wells$Well), "SY (core)", "K (distant)")
cat("Well types:\n")
print(table(wells$Well, wells$type))

# ---- Compute offsets from SY wells only ----
sy_wells <- wells[wells$type == "SY (core)", ]
x0 <- min(sy_wells$X)
y0 <- min(sy_wells$Y)
cat(sprintf("\nOrigin offset (SY min): X0 = %.2f, Y0 = %.2f\n", x0, y0))

# ---- Translate wells_xyz ----
wells$x_local <- wells$X - x0
wells$y_local <- wells$Y - y0
# z downward positive: take absolute values of negative elevations
wells$z_top    <- abs(wells$Z_top)
wells$z_bottom <- abs(wells$Z_bot)

# SY wells summary
cat("\n=== SY well coordinates (translated, z downward positive) ===\n")
sy_wells_t <- wells[wells$type == "SY (core)",
                     c("Well", "Layer", "x_local", "y_local", "z_top", "z_bottom")]
knitr::kable(sy_wells_t, digits = 2,
             caption = "SY1-SY5 translated coordinates")

# Spatial extent of SY wells
cat(sprintf("\nSY wells XY range: x [%.2f, %.2f], y [%.2f, %.2f]\n",
            min(sy_wells$x_local), max(sy_wells$x_local),
            min(sy_wells$y_local), max(sy_wells$y_local)))
cat(sprintf("SY wells Z range: [%.2f, %.2f]\n",
            min(sy_wells$z_top), max(sy_wells$z_bottom)))

# ---- Translate drawdown data ----
drawdown$x_local <- drawdown$X坐标 - x0
drawdown$y_local <- drawdown$Y坐标 - y0

# Parse observation well name and layer
drawdown$obs_well <- gsub("(一层|二层|三层)$", "", drawdown$观测井)
drawdown$obs_layer <- gsub(".*(一层|二层|三层)$", "\\1", drawdown$观测井)
drawdown$layer_num <- ifelse(drawdown$obs_layer == "一层", 1,
                      ifelse(drawdown$obs_layer == "二层", 2, 3))
drawdown$obs_type <- ifelse(grepl("^SY", drawdown$obs_well), "SY (core)", "K (distant)")

cat("\nObservations by well type:\n")
print(table(drawdown$obs_well, drawdown$obs_type))
```

# 4. Merge Pumping Rate Information

```{r merge_pump_rate}
# Merge pump rate into pump_summary
pump_summary <- merge(pump_summary, pump_rate,
                      by.x = "试验名称", by.y = "抽水名",
                      all.x = TRUE)
cat("Pumping test summary with rates:\n")
knitr::kable(pump_summary, digits = 3,
             caption = "Pumping tests with flow rates")
```

# 5. Select 2 Pumping Tests

Select **SY1 Layer1** and **SY2 Layer1** for HT analysis.

```{r select_tests}
selected_tests <- c("SY1一层", "SY2一层")
cat("Selected pumping tests:", selected_tests, "\n\n")

# Filter data
dd_sel <- drawdown[drawdown$试验名称 %in% selected_tests, ]
pump_sel <- pump_summary[pump_summary$试验名称 %in% selected_tests, ]

# Exclude K1-K4 observations: keep only SY observation wells
dd_sel_sy <- dd_sel[dd_sel$obs_type == "SY (core)", ]
cat(sprintf("Total observations: %d -> SY only: %d\n",
            nrow(dd_sel), nrow(dd_sel_sy)))

cat("\nSelected pump tests:\n")
knitr::kable(pump_sel[, c("试验名称", "流量_m3h", "流量_m3min",
                           "持续时间_min", "持续时间_h")],
             digits = 3)
```

# 6. Design Grid with Proper Resolution

Requirements:
- SY1-SY5 wells must be in **different** grid cells
- At least several grid cells as external boundary buffer
- Vertical grid must accurately resolve screen intervals (~7.5m each)
- Minimum well spacing: ~2m (e.g., SY3-SY5 in Y)

Let's compute precisely.

```{r design_grid}
# ---- Analyze well spacing ----
sy_xy <- unique(wells[wells$type == "SY (core)", c("Well", "x_local", "y_local")])
cat("SY well XY positions:\n")
knitr::kable(sy_xy, digits = 2)

# Compute pairwise distances
for (i in 1:(nrow(sy_xy)-1)) {
  for (j in (i+1):nrow(sy_xy)) {
    d <- sqrt((sy_xy$x_local[i] - sy_xy$x_local[j])^2 +
              (sy_xy$y_local[i] - sy_xy$y_local[j])^2)
    if (d < 10) {
      cat(sprintf("  %s - %s: %.2f m\n", sy_xy$Well[i], sy_xy$Well[j], d))
    }
  }
}

# ---- Design grid ----
# SY well XY extent (translated): x ~ [0, 6.8], y ~ [0, 6.6]
# Minimum well spacing: ~2.6 m (SY2-SY5)
#
# Strategy (coarser grid to reduce computation):
#   dx = 1.5 m -> minimum well pair separated by ~2 cells ✓
#   dy = 1.5 m -> minimum well pair separated by ~2 cells ✓
#   Boundary: ~8m (5-6 cells) on each side
#   Total: x: 0-6.8 + 16 margin -> ~23 m (15 cells, dx=1.5)
#          y: 0-6.6 + 16 margin -> ~23 m (15 cells, dy=1.5)
#
# Vertical:
#   Screen intervals: Layer1 [5.6, 13.1], Layer2 [16.6, 24.1], Layer3 [27.6, 35.1]
#   Inter-layer gap: ~3.5m
#   dz = 3.0 m -> each screen spans 2-3 cells, layers separated by ~1 cell ✓
#   Total z: 0 to 40 (14 cells, dz≈2.86m) -> round to dz=3.0, z=[0, 42] (14 cells)

dx_use <- 1.5
dy_use <- 1.5
dz_use <- 3.0

# XY domain with moderate boundary
xy_margin <- 8  # ~5 cells boundary on each side
x1 <- floor(min(sy_xy$x_local)) - xy_margin
x2 <- ceiling(max(sy_xy$x_local)) + xy_margin
y1 <- floor(min(sy_xy$y_local)) - xy_margin
y2 <- ceiling(max(sy_xy$y_local)) + xy_margin

# Ensure integer cell counts
nx <- round((x2 - x1) / dx_use)
ny <- round((y2 - y1) / dy_use)
x2 <- x1 + nx * dx_use  # adjust to exact grid
y2 <- y1 + ny * dy_use

# Z domain
z1 <- 0
z2 <- ceiling(max(wells$z_bottom) + 5)  # 5m below deepest screen
nz <- round((z2 - z1) / dz_use)
z2 <- z1 + nz * dz_use  # adjust to exact grid

domain3d <- c(nx, ny, nz, x1, x2, y1, y2, z1, z2)
cat(sprintf("\nDomain: c(%d, %d, %d, %.1f, %.1f, %.1f, %.1f, %.1f, %.1f)\n",
            nx, ny, nz, x1, x2, y1, y2, z1, z2))

grid3d <- GenGrid3D(domain3d)

cat(sprintf("Grid: %d cells (nx=%d, ny=%d, nz=%d)\n", grid3d$n, nx, ny, nz))
cat(sprintf("Resolution: dx=%.2f, dy=%.2f, dz=%.2f\n",
            grid3d$dx, grid3d$dy, grid3d$dz))

# GenGrid3D does not provide face coordinates; compute them manually
grid3d$xface <- seq(grid3d$x1, grid3d$x2, length.out = grid3d$nx + 1)
grid3d$yface <- seq(grid3d$y1, grid3d$y2, length.out = grid3d$ny + 1)
grid3d$zface <- seq(grid3d$z1, grid3d$z2, length.out = grid3d$nz + 1)

# ---- Check: are SY wells in different cells? ----
cat("\n=== SY well grid cell assignments ===\n")
for (i in 1:nrow(sy_xy)) {
  w <- sy_xy$Well[i]
  ix <- which.min(abs(sy_xy$x_local[i] - grid3d$xmid))
  iy <- which.min(abs(sy_xy$y_local[i] - grid3d$ymid))
  cat(sprintf("  %s: x=%.2f -> cell_ix=%d, y=%.2f -> cell_iy=%d\n",
              w, sy_xy$x_local[i], ix, sy_xy$y_local[i], iy))
}

# Check vertical: screen interval coverage
cat("\n=== Screen interval cell coverage ===\n")
zmid <- grid3d$zmid
for (i in 1:nrow(sy_wells_t)) {
  w  <- sy_wells_t$Well[i]
  ly <- sy_wells_t$Layer[i]
  zt <- sy_wells_t$z_top[i]
  zb <- sy_wells_t$z_bottom[i]
  ncells <- sum(zmid >= zt & zmid <= zb)
  cat(sprintf("  %s %s: z=[%.1f, %.1f] -> %d cells\n", w, ly, zt, zb, ncells))
}
```

# 7. Plot Well Locations on Grid

```{r plot_wells_grid, fig.width=12, fig.height=10}
# Prepare well plot data
well_plot <- wells
well_plot$WellLayer <- paste(well_plot$Well, well_plot$Layer, sep="_")
well_plot$z_mid <- (well_plot$z_top + well_plot$z_bottom) / 2
well_plot$type_label <- ifelse(well_plot$type == "SY (core)", "SY (core)", "K (excluded)")

# XY: wells on grid
p_xy <- ggplot() +
  # Grid lines
  geom_vline(xintercept = grid3d$xface, color = "gray85", linewidth = 0.2) +
  geom_hline(yintercept = grid3d$yface, color = "gray85", linewidth = 0.2) +
  # SY wells (used)
  geom_point(data = well_plot[well_plot$type == "SY (core)", ],
             aes(x = x_local, y = y_local, shape = Layer),
             color = "#E41A1C", size = 3.5) +
  # K wells (excluded)
  geom_point(data = well_plot[well_plot$type == "K (distant)", ],
             aes(x = x_local, y = y_local),
             color = "gray60", size = 2, shape = 4) +
  geom_text(data = well_plot,
            aes(x = x_local, y = y_local, label = WellLayer,
                color = type_label),
            vjust = -1, size = 2.8) +
  scale_color_manual(values = c("SY (core)" = "#E41A1C", "K (excluded)" = "gray60")) +
  scale_shape_manual(values = c("Layer1" = 16, "Layer2" = 17, "Layer3" = 15)) +
  labs(title = "Well Locations with Grid (SY cluster in focus)",
       subtitle = sprintf("Grid: %dx%dx%d, dx=%.1fm, dy=%.1fm, dz=%.1fm",
                          nx, ny, nz, grid3d$dx, grid3d$dy, grid3d$dz),
       x = "X (m, translated)", y = "Y (m, translated)",
       color = "Well Type", shape = "Layer") +
  coord_fixed() +
  theme_minimal()

# XZ: depth profile with grid
p_xz <- ggplot() +
  geom_hline(yintercept = -grid3d$zmid, color = "gray85", linewidth = 0.2) +
  geom_segment(data = well_plot,
               aes(x = x_local, xend = x_local,
                   y = -z_top, yend = -z_bottom,
                   color = type_label),
               linewidth = 2) +
  geom_text(data = well_plot,
            aes(x = x_local, y = -(z_top - 0.5),
                label = WellLayer, color = type_label),
            size = 2.5, angle = 90, hjust = 0) +
  scale_color_manual(values = c("SY (core)" = "#E41A1C", "K (excluded)" = "gray60")) +
  labs(title = "Well Screens with Vertical Grid (XZ view)",
       subtitle = sprintf("dz = %.1f m, %d layers", dz_use, nz),
       x = "X (m)", y = "Depth z (m, negative = deeper)") +
  theme_minimal()

# YZ: depth profile
p_yz <- ggplot() +
  geom_hline(yintercept = -grid3d$zface, color = "gray85", linewidth = 0.2) +
  geom_segment(data = well_plot,
               aes(x = y_local, xend = y_local,
                   y = -z_top, yend = -z_bottom,
                   color = type_label),
               linewidth = 2) +
  geom_text(data = well_plot,
            aes(x = y_local, y = -(z_top - 0.5),
                label = WellLayer, color = type_label),
            size = 2.5, angle = 90, hjust = 0) +
  scale_color_manual(values = c("SY (core)" = "#E41A1C", "K (excluded)" = "gray60")) +
  labs(title = "Well Screens with Vertical Grid (YZ view)",
       x = "Y (m)", y = "Depth z (m)") +
  theme_minimal()

grid.arrange(p_xy, p_xz, p_yz, ncol = 2, nrow = 2,
             layout_matrix = rbind(c(1, 1), c(2, 3)))
```

# 8. Construct qHT and oHT Lists (with Well Screens + Real Qp)

```{r construct_ht}
# ---- Build screen lookup ----
screen_lookup <- wells
screen_lookup$well_layer <- paste(screen_lookup$Well, screen_lookup$Layer, sep="_")

# Build qHT and oHT
qHT <- list()
oHT <- list()
test_info <- list()

for (ti in seq_along(selected_tests)) {
  test_name <- selected_tests[ti]
  cat(sprintf("\n=== Test %d: %s ===\n", ti, test_name))

  # Parse pumping well and layer
  pump_well_name <- gsub("(一层|二层|三层)$", "", test_name)
  pump_layer_str <- gsub(".*(一层|二层|三层)$", "\\1", test_name)
  pump_layer_num <- ifelse(pump_layer_str == "一层", 1,
                    ifelse(pump_layer_str == "二层", 2, 3))

  # ---- Get pump screen info ----
  pump_row <- wells[wells$Well == pump_well_name &
                    wells$Layer == paste0("Layer", pump_layer_num), ]
  if (nrow(pump_row) == 0) {
    cat("  WARNING: pump well not found\n")
    next
  }

  # ---- Get real pumping rate ----
  pr <- pump_summary[pump_summary$试验名称 == test_name, ]
  Qp_real <- if (nrow(pr) > 0 && !is.na(pr$流量_m3min[1])) pr$流量_m3min[1] else 1.0
  cat(sprintf("  Pump rate Qp = %.4f m3/min\n", Qp_real))

  # ---- Build Qinf ----
  Qinf <- data.frame(
    Qp       = Qp_real,
    x        = pump_row$x_local,
    y        = pump_row$y_local,
    z_top    = pump_row$z_top,
    z_bottom = pump_row$z_bottom
  )
  qHT[[test_name]] <- Qinf
  cat(sprintf("  Pump: x=%.2f, y=%.2f, z=[%.2f, %.2f]\n",
              Qinf$x, Qinf$y, Qinf$z_top, Qinf$z_bottom))

  # ---- Build Oinf (SY wells only) ----
  dd_test <- dd_sel_sy[dd_sel_sy$试验名称 == test_name, ]
  cat(sprintf("  SY observation records: %d\n", nrow(dd_test)))

  Oinf_rows <- data.frame(
    data      = -dd_test$最终降深.m.,  # negative: drawdown is below zero-drawdown boundary
    x         = dd_test$x_local,
    y         = dd_test$y_local,
    z_top     = numeric(nrow(dd_test)),
    z_bottom  = numeric(nrow(dd_test)),
    obs_well  = dd_test$obs_well,
    layer_num = dd_test$layer_num,
    stringsAsFactors = FALSE
  )

  # Fill screen coordinates
  for (j in seq_len(nrow(Oinf_rows))) {
    ow <- Oinf_rows$obs_well[j]
    ln <- Oinf_rows$layer_num[j]
    sr <- wells[wells$Well == ow & wells$Layer == paste0("Layer", ln), ]
    if (nrow(sr) > 0) {
      Oinf_rows$z_top[j]    <- sr$z_top
      Oinf_rows$z_bottom[j] <- sr$z_bottom
    } else {
      Oinf_rows$z_top[j]    <- NA
      Oinf_rows$z_bottom[j] <- NA
    }
  }

  Oinf_rows <- Oinf_rows[!is.na(Oinf_rows$z_top), ]
  cat(sprintf("  Valid observations: %d\n", nrow(Oinf_rows)))

  # Print observation summary
  if (nrow(Oinf_rows) > 0) {
    cat("  Observation drawdowns:\n")
    print(summary(Oinf_rows$data))
  }

  Oinf <- Oinf_rows[, c("data", "x", "y", "z_top", "z_bottom")]
  oHT[[test_name]] <- Oinf

  test_info[[test_name]] <- list(
    pump_well  = pump_well_name,
    pump_layer = pump_layer_num,
    Qp         = Qp_real,
    Qinf       = Qinf,
    Oinf       = Oinf
  )
}

cat(sprintf("\n=== Summary ===\n"))
cat(sprintf("Number of pumping tests: %d\n", length(qHT)))
total_obs <- 0
for (nm in names(qHT)) {
  n <- nrow(oHT[[nm]])
  total_obs <- total_obs + n
  cat(sprintf("  %s: Qp=%.4f m3/min, %d observations\n",
              nm, qHT[[nm]]$Qp[1], n))
}
cat(sprintf("Total observations across all tests: %d\n", total_obs))
```

# 9. Drawdown Data Overview

```{r plot_drawdown_overview, fig.width=10, fig.height=6}
# Combine drawdown data for selected tests (SY only)
dd_plot <- dd_sel_sy
dd_plot$obs_well_layer <- paste(dd_plot$obs_well, dd_plot$obs_layer, sep="_")

# Drawdown by observation well
p1 <- ggplot(dd_plot, aes(x = obs_well_layer, y = 最终降深.m., fill = 试验名称)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Final Drawdown by Observation Well",
       x = "Observation Well_Layer", y = "Drawdown (m)") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

# Drawdown map (XY, averaged across layers)
dd_avg <- aggregate(最终降深.m. ~ obs_well + x_local + y_local + 试验名称,
                    data = dd_plot, FUN = mean)
p2 <- ggplot(dd_avg, aes(x = x_local, y = y_local, size = 最终降深.m., color = 试验名称)) +
  geom_point(alpha = 0.7) +
  geom_text(aes(label = obs_well), vjust = -1.5, size = 3) +
  scale_size_continuous(range = c(3, 10)) +
  labs(title = "Drawdown Map (bubble size = drawdown)",
       x = "X (m)", y = "Y (m)") +
  coord_fixed() +
  theme_minimal()

grid.arrange(p1, p2, ncol = 2)
```

# 10. Forward Simulation (Homogeneous K)

Forward simulation to verify the model setup and estimate effective K.

```{r forward_homogeneous}
K_try <- c(0.001, 0.005, 0.01, 0.05, 0.1, 0.5, 1)

forward_results <- list()
for (Ki in K_try) {
  cat(sprintf("\n--- K = %.4f m/min ---\n", Ki))
  for (ti in seq_along(qHT)) {
    test_name <- names(qHT)[ti]
    cat(sprintf("  Test: %s\n", test_name))

    res <- tryCatch({
      Fsteady3dsim(KK = Ki, Qinf = qHT[[ti]], grid = grid3d, lrw = 20000000)
    }, error = function(e) {
      cat(sprintf("    ERROR: %s\n", e$message))
      return(NULL)
    })

    if (!is.null(res)) {
      obs_sampled <- samData3DScreen(Oinf = oHT[[ti]], grid = grid3d,
                                     h = res$solution)
      forward_results[[paste0("K", Ki, "_", test_name)]] <- list(
        result  = res,
        sampled = obs_sampled,
        K       = Ki,
        test    = test_name
      )

      obs_true <- oHT[[ti]]$data
      obs_sim  <- obs_sampled$data
      valid <- !is.na(obs_true) & !is.na(obs_sim) & obs_true < 0 & obs_sim < 0
      if (sum(valid) > 0) {
        ratio <- obs_true[valid] / obs_sim[valid]
        cat(sprintf("    Obs/Sim ratio: mean=%.3f, median=%.3f (n=%d)\n",
                    mean(ratio), median(ratio), sum(valid)))
      }
    }
  }
}
```

# 11. Estimate Effective K from Drawdown Ratios

```{r estimate_K}
# Collect all ratio data
ratio_data <- data.frame()

for (nm in names(forward_results)) {
  fr <- forward_results[[nm]]
  obs_true <- oHT[[fr$test]]$data
  obs_sim  <- fr$sampled$data
  valid <- !is.na(obs_true) & !is.na(obs_sim) & obs_true < 0 & obs_sim < 0
  if (sum(valid) > 0) {
    ratio_data <- rbind(ratio_data, data.frame(
      K = fr$K, ratio = obs_true[valid] / obs_sim[valid],
      test = fr$test, stringsAsFactors = FALSE
    ))
  }
}

if (nrow(ratio_data) > 0) {
  cat("Drawdown ratio (observed/simulated) by K:\n")
  ratio_by_K <- aggregate(ratio ~ K, data = ratio_data,
                          FUN = function(x) c(mean = mean(x), median = median(x)))
  for (i in 1:nrow(ratio_by_K)) {
    cat(sprintf("  K=%.4f: mean_ratio=%.3f, median_ratio=%.3f\n",
                ratio_by_K$K[i],
                ratio_by_K$ratio[i, "mean"],
                ratio_by_K$ratio[i, "median"]))
  }

  # Estimate K where ratio ≈ 1
  ratio_med <- aggregate(ratio ~ K, data = ratio_data, FUN = median)
  names(ratio_med)[2] <- "median_ratio"

  # Linear interpolation in log-log space
  if (min(ratio_med$median_ratio) < 1 && max(ratio_med$median_ratio) > 1) {
    lo <- max(which(ratio_med$median_ratio < 1))
    hi <- min(which(ratio_med$median_ratio > 1))
    K_est <- 10^(log10(ratio_med$K[lo]) +
                 (log10(1) - log10(ratio_med$median_ratio[lo])) /
                 (log10(ratio_med$median_ratio[hi]) -
                  log10(ratio_med$median_ratio[lo])) *
                 (log10(ratio_med$K[hi]) - log10(ratio_med$K[lo])))
  } else {
    K_est <- ratio_med$K[which.min(abs(ratio_med$median_ratio - 1))]
  }
  cat(sprintf("\n=== Estimated effective K: %.6f m/min ===\n", K_est))

  # Plot ratio vs K
  p_ratio <- ggplot(ratio_data, aes(x = factor(K), y = ratio, fill = test)) +
    geom_boxplot(alpha = 0.6) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "red", linewidth = 1) +
    scale_y_log10() +
    labs(title = "Observed/Simulated Drawdown Ratio vs K",
         subtitle = sprintf("Estimated K = %.6f m/min", K_est),
         x = "K (m/min)", y = "Ratio (observed/simulated)") +
    theme_minimal()
  print(p_ratio)
} else {
  K_est <- 0.01
  cat("No valid ratio data. Using default K = 0.01\n")
}
```

# 12. 1:1 Plot: Observed vs Simulated (Homogeneous Forward)

```{r plot_obs_vs_sim, fig.width=10, fig.height=7}
plot_data <- data.frame()
for (nm in names(forward_results)) {
  fr <- forward_results[[nm]]
  obs <- oHT[[fr$test]]$data
  sim <- fr$sampled$data
  plot_data <- rbind(plot_data, data.frame(
    observed = obs, simulated = sim,
    K = fr$K, test = fr$test,
    stringsAsFactors = FALSE
  ))
}

# Both observed and simulated are negative (drawdown); use absolute for log-log
plot_data <- plot_data[plot_data$observed < 0 & plot_data$simulated < 0, ]
plot_data$observed_abs  <- abs(plot_data$observed)
plot_data$simulated_abs <- abs(plot_data$simulated)

if (nrow(plot_data) > 0) {
  # Find the K closest to K_est for highlighting
  best_K <- K_try[which.min(abs(K_try - K_est))]
  plot_data$best <- plot_data$K == best_K

  p <- ggplot(plot_data, aes(x = observed_abs, y = simulated_abs)) +
    geom_point(aes(color = factor(K), alpha = best), size = 2.5) +
    geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "gray50") +
    scale_alpha_manual(values = c("TRUE" = 1, "FALSE" = 0.3), guide = "none") +
    scale_x_log10() + scale_y_log10() +
    facet_wrap(~test, scales = "free") +
    labs(title = "Observed vs Simulated Drawdown (Homogeneous K)",
         subtitle = sprintf("Estimated K = %.6f, best trial K = %.4f", K_est, best_K),
         x = "|Observed| (m)", y = "|Simulated| (m)", color = "K") +
    theme_minimal()
  print(p)
}
```

# 13. Forward Simulation Drawdown Field

```{r plot_forward_field, fig.width=12, fig.height=10}
# Plot drawdown field for the best K estimate
if (exists("K_est") && !is.na(K_est)) {
  for (ti in seq_along(qHT)) {
    test_name <- names(qHT)[ti]
    cat(sprintf("\nPlotting drawdown field: %s, K=%.6f\n", test_name, K_est))

    res_fwd <- Fsteady3dsim(KK = K_est, Qinf = qHT[[ti]], grid = grid3d,
                            lrw = 20000000)

    # Sample observations
    obs_sampled <- samData3DScreen(Oinf = oHT[[ti]], grid = grid3d,
                                   h = res_fwd$solution)

    # Select 3 z-slices: near each layer center
    z_layers <- unique(grid3d$zmid)
    z_targets <- c(9, 20, 31)  # approximate centers of 3 screen layers
    z_sel <- sapply(z_targets, function(zt) z_layers[which.min(abs(z_layers - zt))])
    z_sel <- unique(z_sel)

    for (z_val in z_sel) {
      slice_data <- res_fwd[abs(res_fwd$z - z_val) < 1e-10, ]

      # Add observation wells at this depth layer
      obs_at_z <- obs_sampled
      # Find which layer this z belongs to
      layer_label <- if (z_val < 14) "~Layer1" else if (z_val < 26) "~Layer2" else "~Layer3"

      p <- ggplot(slice_data, aes(x = x, y = y, fill = solution)) +
        geom_tile() +
        scale_fill_viridis_c(option = "plasma", name = "Drawdown (m)") +
        # Pump location
        geom_point(data = qHT[[ti]],
                   aes(x = x, y = y), color = "red", size = 3, shape = 8,
                   inherit.aes = FALSE) +
        # SY observation wells
        geom_point(data = wells[wells$type == "SY (core)", ],
                   aes(x = x_local, y = y_local),
                   color = "white", size = 1.5, alpha = 0.7,
                   inherit.aes = FALSE) +
        labs(title = sprintf("Drawdown Field: %s, z=%.1f m (%s)",
                             test_name, z_val, layer_label),
             x = "X (m)", y = "Y (m)") +
        coord_fixed() +
        theme_minimal()
      print(p)
    }
  }
}
```

# 14. Adjoint-Based Inversion (Steady-State 3D with Well Screens)

This section uses the **adjoint-state method** for computing the Jacobian (sensitivity) matrix analytically, replacing the ensemble-based approach. The adjoint method is deterministic and typically faster for problems with moderate grid sizes.

Key differences from the ensemble version:
- No ensemble / Monte Carlo sampling; uses analytical Jacobian
- Levenberg-Marquardt damping for regularization
- Controlled by `sigma2obs` (observation error variance), `lambda_lm` (LM damping), and `decay_lm` (damping decay factor)

```{r inversion_adj}
# Use estimated K as prior geometric mean
prior_K <- if (exists("K_est") && !is.na(K_est) && K_est > 0) K_est else 0.01

cat(sprintf("Prior geometric mean K = %.6f m/min\n", prior_K))

# Geo-statistical parameters
geo_params <- list(
  me     = log(prior_K),
  var    = 1.5,
  geomod = "Exp",
  range  = 15,
  nugget = 0.,
  anis   = c(0, 0, 0, 1, 0.2)
)

# Adjoint inversion parameters
itermax <- 20
sigma2obs <- 1e-4      # observation error variance
lambda_lm <- 1.0        # initial Levenberg-Marquardt damping
decay_lm  <- 1.4        # LM damping decay factor

cat(sprintf("Max iterations: %d\n", itermax))
cat(sprintf("sigma2obs = %.1e, lambda_lm = %.1f, decay_lm = %.1f\n",
            sigma2obs, lambda_lm, decay_lm))
cat(sprintf("Grid: %d cells, %d observations\n", grid3d$n,
            sum(sapply(oHT, nrow))))

# Run adjoint-based inversion
inv_result_adj <- Finverse3DScreenAdj(
  grid        = grid3d,
  qHT         = qHT,
  oHT         = oHT,
  itermax     = itermax,
  rmsemin     = 0,
  lrw         = 20000000,
  sigma2obs   = sigma2obs,
  lambda_lm   = lambda_lm,
  decay_lm    = decay_lm,
  geo         = geo_params,
  ifcor       = FALSE,
  ifvarTh     = FALSE
)

cat("\n=== Adjoint inversion completed! ===\n")
cat(sprintf("Number of iterations: %d\n", length(inv_result_adj)))
```

# 15. Inversion Results: Convergence and K Field

```{r inv_results_adj, fig.width=14, fig.height=12}
niter_done <- length(inv_result_adj)
cat(sprintf("Number of iterations: %d\n", niter_done))
niter_done = 12
# ---- RMSE convergence ----
rmse_vec <- sapply(inv_result_adj, function(x) x$rmse)
cat("\n=== RMSE Convergence ===\n")
for (i in seq_along(rmse_vec)) {
  cat(sprintf("  Iter %d: RMSE = %.6f\n", i, rmse_vec[i]))
}

rmse_df <- data.frame(iteration = seq_along(rmse_vec), RMSE = rmse_vec)
p_rmse <- ggplot(rmse_df, aes(x = iteration, y = RMSE)) +
  geom_line(color = "#377EB8", linewidth = 1) +
  geom_point(size = 4, color = "#E41A1C") +
  labs(title = "Convergence: RMSE (Adjoint Method)",
       x = "Iteration", y = "RMSE") +
  theme_minimal()
print(p_rmse)

# ---- Use inversePlot3D for last iteration ----
figs <- inversePlot3D(
  niterm = niter_done,
  grid   = grid3d,
  iterdf = inv_result_adj,
  oHT    = oHT,
  zslice = grid3d$zmid[round(grid3d$nz / 2)]
)

# Add well locations to ln(K) and variance slices
sy_wells_plot <- wells[wells$type == "SY (core)", ]
figs$lnk <- figs$lnk +
  geom_point(data = sy_wells_plot,
             aes(x = x_local, y = y_local),
             color = "white", size = 2, shape = 21, fill = NA,
             stroke = 0.5, inherit.aes = FALSE) +
  geom_text(data = sy_wells_plot,
            aes(x = x_local, y = y_local, label = Well),
            color = "white", size = 2.5, vjust = -1,
            inherit.aes = FALSE)

if (!is.null(figs$varlnk)) {
  figs$varlnk <- figs$varlnk +
    geom_point(data = sy_wells_plot,
               aes(x = x_local, y = y_local),
               color = "white", size = 2, shape = 21, fill = NA,
               stroke = 0.5, inherit.aes = FALSE) +
    geom_text(data = sy_wells_plot,
              aes(x = x_local, y = y_local, label = Well),
              color = "white", size = 2.5, vjust = -1,
              inherit.aes = FALSE)
}

print(ggpubr::ggarrange(plotlist = figs, ncol = 2, nrow = 2))

# ---- K statistics ----
last_iter <- inv_result_adj[[niter_done]]
Kmean <- exp(last_iter$meanT)
cat(sprintf("\nK statistics: mean=%.6f, median=%.6f, sd=%.6f\n",
            mean(Kmean), median(Kmean), sd(Kmean)))
cat(sprintf("K range: [%.6f, %.6f]\n", min(Kmean), max(Kmean)))
```

# 16. Final Comparison: Observed vs Predicted

```{r final_comparison_adj, fig.width=10, fig.height=7}
K_mean_final <- exp(last_iter$meanT)  # meanT is mean of ln(K)

compare_data <- data.frame()
for (ti in seq_along(qHT)) {
  test_name <- names(qHT)[ti]

  res_final <- Fsteady3dsim(KK = K_mean_final, Qinf = qHT[[ti]], grid = grid3d,
                             lrw = 20000000)

  obs_sampled <- samData3DScreen(Oinf = oHT[[ti]], grid = grid3d,
                                 h = res_final$solution)
  compare_data <- rbind(compare_data, data.frame(
    observed  = oHT[[ti]]$data,
    predicted = obs_sampled$data,
    test      = test_name,
    stringsAsFactors = FALSE
  ))
}

compare_data <- compare_data[compare_data$observed < 0 &
                             compare_data$predicted < 0, ]
compare_data$observed_abs  <- abs(compare_data$observed)
compare_data$predicted_abs <- abs(compare_data$predicted)

# Compute R²
fit <- lm(predicted ~ observed, data = compare_data)
r2 <- summary(fit)$r.squared
rmse_val <- sqrt(mean((compare_data$observed - compare_data$predicted)^2))

p <- ggplot(compare_data, aes(x = observed_abs, y = predicted_abs, color = test)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "gray50", linewidth = 0.8) +
  geom_smooth(method = "lm", se = TRUE, aes(group = 1),
              color = "#377EB8", linewidth = 0.8, alpha = 0.2) +
  annotate("text",
           x = max(compare_data$observed_abs) * 0.4,
           y = max(compare_data$predicted_abs) * 0.9,
           label = sprintf("R² = %.3f\nRMSE = %.4f m", r2, rmse_val),
           size = 5, hjust = 0) +
  scale_x_log10() + scale_y_log10() +
  labs(title = "Final Comparison: Observed vs Predicted Drawdown (Adjoint)",
       subtitle = "Using adjoint-inverted heterogeneous K field",
       x = "|Observed| Drawdown (m)",
       y = "Predicted Drawdown (m)") +
  theme_minimal()
print(p)

cat(sprintf("\n=== Final Statistics (Adjoint) ===\n"))
cat(sprintf("R-squared: %.4f\n", r2))
cat(sprintf("RMSE: %.6f m\n", rmse_val))

# Per-test statistics
for (tn in unique(compare_data$test)) {
  cd <- compare_data[compare_data$test == tn, ]
  cat(sprintf("  %s: n=%d, R²=%.4f, RMSE=%.6f\n",
              tn, nrow(cd),
              summary(lm(predicted ~ observed, data = cd))$r.squared,
              sqrt(mean((cd$observed - cd$predicted)^2))))
}
```

# 17. Refined Inversion: z-direction Doubled, Adjusted Geo-Parameters

```{r refined_inversion_adj}
cat("=== Building refined grid (z doubled) ===\n")

# z-direction: double the resolution (half dz)
dz_refined <- dz_use / 2  # e.g. 3.0 -> 1.5
nz_refined <- nz * 2
z2_refined <- z1 + nz_refined * dz_refined

domain3d_refined <- c(nx, ny, nz_refined, x1, x2, y1, y2, z1, z2_refined)
cat(sprintf("Domain: c(%d, %d, %d, %.1f, %.1f, %.1f, %.1f, %.1f, %.1f)\n",
            nx, ny, nz_refined, x1, x2, y1, y2, z1, z2_refined))

grid3d_r <- GenGrid3D(domain3d_refined)

cat(sprintf("Grid: %d cells (nx=%d, ny=%d, nz=%d)\n", grid3d_r$n, nx, ny, nz_refined))
cat(sprintf("Resolution: dx=%.2f, dy=%.2f, dz=%.2f\n",
            grid3d_r$dx, grid3d_r$dy, grid3d_r$dz))

# Face coordinates
grid3d_r$xface <- seq(grid3d_r$x1, grid3d_r$x2, length.out = grid3d_r$nx + 1)
grid3d_r$yface <- seq(grid3d_r$y1, grid3d_r$y2, length.out = grid3d_r$ny + 1)
grid3d_r$zface <- seq(grid3d_r$z1, grid3d_r$z2, length.out = grid3d_r$nz + 1)

# ---- Check screen coverage on refined grid ----
cat("\n=== Screen interval cell coverage (refined z) ===\n")
zmid_r <- grid3d_r$zmid
for (i in 1:nrow(sy_wells_t)) {
  w  <- sy_wells_t$Well[i]
  ly <- sy_wells_t$Layer[i]
  zt <- sy_wells_t$z_top[i]
  zb <- sy_wells_t$z_bottom[i]
  ncells <- sum(zmid_r >= zt & zmid_r <= zb)
  cat(sprintf("  %s %s: z=[%.1f, %.1f] -> %d cells\n", w, ly, zt, zb, ncells))
}

# ---- Adjusted geo-parameters: xy range=25, z range=3 ----
geo_params_refined <- list(
  me     = log(prior_K),
  var    = 1.5,
  geomod = "Exp",
  range  = 25,                          # major range (x direction)
  nugget = 0.1,
  anis   = c(0, 0, 0, 1, 3/25)         # ratio_y=1, ratio_z=3/25=0.12
)

cat(sprintf("\nGeo-params: xy_range=25, z_range=3\n"))

# ---- Run refined adjoint inversion ----
inv_result_r_adj <- Finverse3DScreenAdj(
  grid        = grid3d_r,
  qHT         = qHT,
  oHT         = oHT,
  itermax     = itermax,
  rmsemin     = 0,
  lrw         = 20000000,
  sigma2obs   = sigma2obs,
  lambda_lm   = lambda_lm,
  decay_lm    = decay_lm,
  geo         = geo_params_refined,
  ifcor       = FALSE,
  ifvarTh     = FALSE
)

cat("\n=== Refined adjoint inversion completed! ===\n")
cat(sprintf("Number of iterations: %d\n", length(inv_result_r_adj)))
```

```{r refined_results_adj, fig.width=14, fig.height=12}
niter_r <- length(inv_result_r_adj)
cat(sprintf("Number of iterations: %d\n", niter_r))

# ---- RMSE convergence ----
rmse_vec_r <- sapply(inv_result_r_adj, function(x) x$rmse)
cat("\n=== RMSE Convergence (Refined) ===\n")
for (i in seq_along(rmse_vec_r)) {
  cat(sprintf("  Iter %d: RMSE = %.6f\n", i, rmse_vec_r[i]))
}

rmse_df_r <- data.frame(iteration = seq_along(rmse_vec_r), RMSE = rmse_vec_r)
p_rmse_r <- ggplot(rmse_df_r, aes(x = iteration, y = RMSE)) +
  geom_line(color = "#377EB8", linewidth = 1) +
  geom_point(size = 4, color = "#E41A1C") +
  labs(title = "Convergence: RMSE — Refined Grid (Adjoint)",
       x = "Iteration", y = "RMSE") +
  theme_minimal()
print(p_rmse_r)

# ---- inversePlot3D ----
figs_r <- inversePlot3D(
  niterm = niter_r,
  grid   = grid3d_r,
  iterdf = inv_result_r_adj,
  oHT    = oHT,
  zslice = grid3d_r$zmid[round(grid3d_r$nz / 2)]
)

figs_r$lnk <- figs_r$lnk +
  geom_point(data = sy_wells_plot,
             aes(x = x_local, y = y_local),
             color = "white", size = 2, shape = 21, fill = NA,
             stroke = 0.5, inherit.aes = FALSE) +
  geom_text(data = sy_wells_plot,
            aes(x = x_local, y = y_local, label = Well),
            color = "white", size = 2.5, vjust = -1,
            inherit.aes = FALSE)

if (!is.null(figs_r$varlnk)) {
  figs_r$varlnk <- figs_r$varlnk +
    geom_point(data = sy_wells_plot,
               aes(x = x_local, y = y_local),
               color = "white", size = 2, shape = 21, fill = NA,
               stroke = 0.5, inherit.aes = FALSE) +
    geom_text(data = sy_wells_plot,
              aes(x = x_local, y = y_local, label = Well),
              color = "white", size = 2.5, vjust = -1,
              inherit.aes = FALSE)
}

print(ggpubr::ggarrange(plotlist = figs_r, ncol = 2, nrow = 2))

# ---- K statistics ----
last_iter_r <- inv_result_r_adj[[niter_r]]
Kmean_r <- exp(last_iter_r$meanT)
cat(sprintf("\nK statistics (refined): mean=%.6f, median=%.6f, sd=%.6f\n",
            mean(Kmean_r), median(Kmean_r), sd(Kmean_r)))
cat(sprintf("K range: [%.6f, %.6f]\n", min(Kmean_r), max(Kmean_r)))
```

```{r refined_comparison_adj, fig.width=10, fig.height=7}
# Observed vs Predicted for refined adjoint inversion
compare_data_r <- data.frame()
for (ti in seq_along(qHT)) {
  test_name <- names(qHT)[ti]

  res_final_r <- Fsteady3dsim(KK = Kmean_r, Qinf = qHT[[ti]], grid = grid3d_r,
                               lrw = 20000000)

  obs_sampled_r <- samData3DScreen(Oinf = oHT[[ti]], grid = grid3d_r,
                                   h = res_final_r$solution)
  compare_data_r <- rbind(compare_data_r, data.frame(
    observed  = oHT[[ti]]$data,
    predicted = obs_sampled_r$data,
    test      = test_name,
    stringsAsFactors = FALSE
  ))
}

compare_data_r <- compare_data_r[compare_data_r$observed < 0 &
                                 compare_data_r$predicted < 0, ]
compare_data_r$observed_abs  <- abs(compare_data_r$observed)
compare_data_r$predicted_abs <- abs(compare_data_r$predicted)

fit_r <- lm(predicted ~ observed, data = compare_data_r)
r2_r <- summary(fit_r)$r.squared
rmse_r <- sqrt(mean((compare_data_r$observed - compare_data_r$predicted)^2))

p_r <- ggplot(compare_data_r, aes(x = observed_abs, y = predicted_abs, color = test)) +
  geom_point(size = 3, alpha = 0.7) +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed",
              color = "gray50", linewidth = 0.8) +
  geom_smooth(method = "lm", se = TRUE, aes(group = 1),
              color = "#377EB8", linewidth = 0.8, alpha = 0.2) +
  annotate("text",
           x = max(compare_data_r$observed_abs) * 0.4,
           y = max(compare_data_r$predicted_abs) * 0.9,
           label = sprintf("R² = %.3f\nRMSE = %.4f m", r2_r, rmse_r),
           size = 5, hjust = 0) +
  scale_x_log10() + scale_y_log10() +
  labs(title = "Refined Adjoint Inversion: Observed vs Predicted Drawdown",
       subtitle = "z-doubled grid, xy_range=25, z_range=3",
       x = "|Observed| Drawdown (m)",
       y = "Predicted Drawdown (m)") +
  theme_minimal()
print(p_r)

cat(sprintf("\n=== Refined Final Statistics (Adjoint) ===\n"))
cat(sprintf("R-squared: %.4f\n", r2_r))
cat(sprintf("RMSE: %.6f m\n", rmse_r))
```

# 18. K Field Cross-Section (XZ and YZ Profiles)

```{r xz_profile_adj, fig.width=12, fig.height=6}
K_grid <- data.frame(grid3d$grid, K = Kmean)

# XZ slice at y ~ middle of SY cluster
y_mid <- mean(range(wells$y_local[wells$type == "SY (core)"]))
y_idx <- which.min(abs(grid3d$ymid - y_mid))
y_val <- grid3d$ymid[y_idx]

K_xz <- K_grid[abs(K_grid$y - y_val) < 1e-10, ]
K_xz$depth <- -K_xz$z  # negative for display

p_xz <- ggplot(K_xz, aes(x = x, y = depth, fill = log10(K))) +
  geom_tile() +
  scale_fill_viridis_c(option = "viridis", name = "log10(K)") +
  # Overlay screen intervals projected onto this Y slice
  geom_segment(data = wells[wells$type == "SY (core)" &
                            abs(wells$y_local - y_val) < 3, ],
               aes(x = x_local, xend = x_local,
                   y = -z_top, yend = -z_bottom),
               color = "white", linewidth = 1.5, inherit.aes = FALSE) +
  labs(title = sprintf("K Field XZ Cross-Section — Adjoint (y = %.1f m)", y_val),
       x = "X (m)", y = "Depth (m)") +
  theme_minimal()
print(p_xz)

# YZ slice at x ~ middle
x_mid <- mean(range(wells$x_local[wells$type == "SY (core)"]))
x_idx <- which.min(abs(grid3d$xmid - x_mid))
x_val <- grid3d$xmid[x_idx]

K_yz <- K_grid[abs(K_grid$x - x_val) < 1e-10, ]
K_yz$depth <- -K_yz$z

p_yz <- ggplot(K_yz, aes(x = y, y = depth, fill = log10(K))) +
  geom_tile() +
  scale_fill_viridis_c(option = "viridis", name = "log10(K)") +
  geom_segment(data = wells[wells$type == "SY (core)" &
                            abs(wells$x_local - x_val) < 3, ],
               aes(x = y_local, xend = y_local,
                   y = -z_top, yend = -z_bottom),
               color = "white", linewidth = 1.5, inherit.aes = FALSE) +
  labs(title = sprintf("K Field YZ Cross-Section — Adjoint (x = %.1f m)", x_val),
       x = "Y (m)", y = "Depth (m)") +
  theme_minimal()
print(p_yz)

```

# Session Info

```{r session_info}
sessionInfo()
```
