---
title: "HydroTomo"
author:
  - name: "Yuanyuan Zha"
    affiliation: "Wuhan University"
    email: "zhayuan87@whu.edu.cn"
    orcid: "0000-0003-4323-0730"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{HydroTomo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width=8,
  fig.height=6
)
```

# Installation

You can install the development version from GitHub:

```{r install, eval=FALSE}
# install.packages("devtools")
# devtools::install_github("zhayuan87/HydroTomo")
```

# Package structure

- **Forward modeling** — `Fsteady2dsim()`, `Fsteady3dsim()`, `Fsteady2dtrsim()`
  *Simulate steady-state drawdown in 2D/3D confined aquifers under pumping tests.*

- **Random field generation** — `random2d()`, `random3d()`
  *Generate random hydraulic conductivity/transmissivity fields via geostatistical simulation.*

- **Sampling** — `samData()`
  *Extract simulated drawdown at observation well locations.*

- **Sensitivity analysis** — `jacobian2D()`, `jacobian3D()`
  *Compute Jacobian (sensitivity) matrices via the adjoint-state method.*

- **Inversion** — `Finverse()`, `FinverseAdj()`, `Finverse2D()`, `Finverse3D()`
  *Perform HT inversion using geostatistical or adjoint-based approaches.*

- **Ensemble runs** — `samHT()`, `samHT3D()`, `samHTtr()`, `samHTmc()`, `samHTmcPar()`, `samHTmcPar3D()`
  *Run forward simulations for a single realization or an ensemble (sequential or parallel).*

- **Visualization** — `Plotsteady2d()`, `Plotsteady3d()`, `covhkPlot2D()`, `inversePlot()`, `HydroPlot()`
  *Plot drawdown fields, cross-covariance maps, and inversion results.*

# Background 

This package is developed for doing Hydraulic Tomography (HT) inversion in R. The HT technique or philosophy is to conduct multiple pumping tests in a confined aquifer and use the drawdown data to estimate the heterogeneous hydraulic conductivity field. The package contains functions for generating random hydraulic conductivity fields, simulating steady-state groundwater flow under pumping test conditions, performing HT inversion using geostatistical approach, and visualizing the results. 

The developer, Yuanyuan Zha, was a visiting scholar and postdoc at University of Arizona, supervised by Professor Jim Yeh, who is the origional developer of the HT method. The package is mainly based on the codes and algorithms developed during Yuanyuan's staying at Wuhan University.

# Example of toy forward studies

In the package, there is some default parameters in the function to make the simulation and inversion easier. Here is an example of simulating steady-state drawdown data for a pumping test in a confined aquifer with a default homogeneous transmissivity field.

```{r toyforward}
library('HydroTomo')
s <- Fsteady2dsim()
Plotsteady2d(s)
Qinf = data.frame(Qp=c(10,10),x=c(10.5,30.5),y=c(20.5,20.5))
s <- Fsteady2dsim(Qinf=Qinf)
Plotsteady2d(s)
```

We also give example of simulating steady-state drawdown data with heterogeneous transmissivity field.

```{r toyforward2}
set.seed(100)
TT <- random2d(nsim=1)$Tp
Qinf = data.frame(Qp=c(10,10),x=c(10.5,30.5),y=c(20.5,20.5))
s <- Fsteady2dsim(TT=TT,Qinf=Qinf)
Plotsteady2d(s)
```

# Example of toy inverse studies

Now we give a simple example of inverse method. Since it is very time consuming to do a full inversion, here we just show how to run the function with small number of iterations and small ensemble size.

```{r toyinverse,cache=TRUE}
set.seed(100)
trueK <- random2d(nsim=1)
TT <- trueK[,-c(1,2)]
domain=c(40,40,0,40,0,40)
grid = GenGrid(domain)
Qinf1=data.frame(Qp=10,xp=20.5,yp=20.5)
qHT <- list(test1 = Qinf1)
trueh <- Fsteady2dsim(TT=TT,Qinf=Qinf1,domain=domain)
locx = c(15,18,22,25,30)
locy = c(15,18,22,25,30)
loc= expand.grid(x=locx,y=locy)
Oinf1 <- data.frame(data=NA,x=loc$x,y=loc$y)
Oinf1 <- samData(grid = grid,Oinf = Oinf1,h=trueh$solution)
oHT <- list(test1 = Oinf1)
result <- Finverse(grid =grid, qHT = qHT, oHT = oHT)
inversePlot(niterm=5,grid=grid,iterdf = result,oHT = oHT,trueK=trueK)
```
