---
title: "hydraulic tomography incremental paper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hydraulic tomography incremental paper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is to create the paper data and figures.

```{r setup}
library(HydroTomo)
```

# Fig. 1

This figure draw the synthetic case of hydraulic tomography. In figure a, we generate a random field, and gives the 6 pumping well locations. In figure b, we show the simulated drawdown data for test A-1 and also the location of observation wells in test A-1.

```{r figure1}
set.seed(2025)
data('qHT')
library(dplyr)
qHTdf = dplyr::bind_rows(qHT,.id='id')
qHTdf = qHTdf %>% dplyr::filter(id %in% c('A1-1','A2-1','A3-1','A4-1','A5-1','A6-1'))

domain=c(80,80,0,80,0,80)
grid <- GenGrid(domain=domain)
# exp(3.5)=33.12 m2/d means very large T values in sand and gravel aquifer.
geo = list(me = 3.2, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0)
trueK <- random2d(nsim=1,grid=grid,geo = geo)
trueK$Tp = log(trueK$Tp)
# make sure chinese font show correctly
library(ggplot2)
library(showtext)
showtext_auto()
  p1 = ggplot() +
    geom_tile(data = trueK, aes(x, y, fill = Tp)) +
    scale_fill_viridis_c(option = "D")+
    geom_point(data = qHTdf, aes(x = x, y = y), color = "white", size = 3, shape=17)+
   geom_text(data = qHTdf, aes(x = x, y = y, label = id), vjust = -1, color = "black")+
    labs(title = "(a) 参考lnT及抽水井", x = "X/m", y = "Y/m", fill = "log(T)")+
    theme_minimal()
  p1
```

```{r figure1b}
# simulate drawdown data for test A1-1
TT <- exp(trueK$Tp)
Qinf = qHT[['A1-1']]
data('oHT')
library(dplyr)
oHTdf = dplyr::bind_rows(oHT,.id='id')
oHTdf = oHTdf %>% dplyr::filter(id %in% c('A1-1'))
s = Fsteady2dsim(TT=TT,Qinf=Qinf,grid=grid,lrw=320000)

  p2 = ggplot() +
    geom_tile(data = s, aes(x, y, fill = -solution)) +
    scale_fill_viridis_c(option = "D")+
    geom_point(data = oHTdf, aes(x = x, y = y), color = "red",fill="white", size = 3, shape=21)+
    labs(title = "(b) 模拟降深及观测井", x = "X/m", y = "Y/m", fill = "降深/m")+
    theme_minimal()
  p2
```  
Now we use ggarrange to make the p1 p2 together.

```{r figure1c, fig.width=10, fig.height=5}
library(ggpubr)

p = ggarrange(p1, p2, ncol = 2, nrow = 1)

p
```

# Generating synthetic data

We then simulate steady-state drawdown data for a pumping test using the true transmissivity field.

```{r generate_synthetic_data, eval=FALSE}
# TT <- 
qHT1 = qHT[c(1,3,5,7,9,11)]
oHT1 = oHT[c(1,3,5,7,9,11)]
### loop over the list to update the value in oHT1.

for (i in 1:length(oHT1)){
  Qinf <- qHT1[[i]]
  trueh <- Fsteady2dsim(TT=TT,Qinf=Qinf,grid=grid,lrw=320000)$solution
  oinf <- oHT1[[i]]
  oinf = samData(grid = grid,Oinf = oinf,h=trueh)
  oHT1[[i]] <- oinf[,c('data','x','y')]
}

dfoHT1 = do.call(rbind,oHT1)
mean_oHT1 = mean(dfoHT1$data)
print(paste("The mean value of oHT1 is ", round(mean_oHT1,4)))
```

save the best synthetic data.

```{r save_data, eval=FALSE}
blh_syn_best_truth <- list(trueK=trueK, qHT=qHT1, oHT=oHT1, grid=grid)
usethis::use_data(blh_syn_best_truth, overwrite = TRUE)
```
Now we use "blh_synthetic" data to do a sequence of inversion and save the result.

```{r perform_inversion, cache=TRUE, eval=FALSE}
data('blh_synthetic')
grid = blh_synthetic$grid
lire = list()
# we select qHT1 and oHT1 from blh_synthetic, first test1, then 1,2, then 1,2,3
for(i in 1:6){
  qHT1 = blh_synthetic$qHT[1:i]
  oHT1 = blh_synthetic$oHT[1:i]
  result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT1, nsim=50, itermax=10, lrw=320000,
                     geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0))
  lire[[i]] = result
}
### save the result
saveRDS(lire, file = "lire_blh_synthetic_no-obs-error.rds")
```

# Plotting inversion results

```{r plot_inversion_results, eval=TRUE}
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
lire = readRDS(file = "E:/HydroTomo/lire_blh_synthetic_no-obs-error.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = paste0("(",letters[1:6],") " ,1:6,"次抽水试验")
for(i in 1:6){
  niterm = length(lire[[i]])
  oHT1 = blh_synthetic$oHT[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m",
                    p4title = i1[i],
                    p4xlab = "参考lnT",
                    p4ylab = "估计lnT")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol = 3, nrow = 2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol = 3, nrow = 2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol = 3, nrow = 2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol = 3, nrow = 2)
p4
```

# noised data

The head match perfectly, but in reality the head data often contains noise. We can add noise to the synthetic data and perform inversion again.

```{r add_noise_and_invert, cache=TRUE, eval=FALSE}
set.seed(300)
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
qHT1 = blh_synthetic$qHT
oHT1 = blh_synthetic$oHT
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}

lire = list()
# we select qHT1 and oHT1 from blh_synthetic, first test1, then 1,2, then 1,2,3
for(i in 1:6){
  qHT1 = blh_synthetic$qHT[1:i]
  oHT1 = oHT_noised[1:i]
  result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT1, nsim=50, itermax=10, lrw=320000,
                     geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0))
  lire[[i]] = result
}
### save the result
saveRDS(lire, file = "result_blh_synthetic_with-obs-error-sd0_05.rds")
```



```{r increase_noise_and_invert, cache=TRUE, eval=FALSE}
set.seed(400)
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
qHT1 = blh_synthetic$qHT
oHT1 = blh_synthetic$oHT
oHT_noised2 <- oHT1
for (i in 1:length(oHT_noised2)){
  noise = rnorm(n=nrow(oHT_noised2[[i]]), mean=0, sd=0.2*sd(oHT_noised2[[i]]$data))
  oHT_noised2[[i]]$data <- oHT_noised2[[i]]$data + noise
}
lire = list()
# we select qHT1 and oHT1 from blh_synthetic, first test1, then 1,2, then 1,2,3
for(i in 1:6){
  qHT1 = blh_synthetic$qHT[1:i]
  oHT1 = oHT_noised2[1:i]
  result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT1, nsim=50, itermax=10, lrw=320000,
                     geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0))
  lire[[i]] = result
}
### save the result
saveRDS(lire, file = "result_blh_synthetic_with-obs-error-sd0_2.rds")
```

# plot noised data inversion results

```{r plot_noised_inversion_results, eval=FALSE}
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_05.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = paste0("(",letters[1:6],") " ,1:6,"次抽水试验，观测数据含5%噪声")
for(i in 1:6){
  niterm = length(lire[[i]])
  oHT1 = blh_synthetic$oHT[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m",
                    p4title = i1[i],
                    p4xlab = "参考lnT",
                    p4ylab = "估计lnT")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk 
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol = 3, nrow = 2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol = 3, nrow = 2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol = 3, nrow = 2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol = 3, nrow = 2)
p4
```

# more noise inversion results
```{r plot_more_noised_inversion_results, eval=FALSE}
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_2.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = paste0("(",letters[1:6],") " ,1:6,"次抽水试验，观测数据含20%噪声")
for(i in 1:6){
  niterm = length(lire[[i]])
  oHT1 = blh_synthetic$oHT[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m",
                    p4title = i1[i],
                    p4xlab = "参考lnT",
                    p4ylab = "估计lnT")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol = 3, nrow = 2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol = 3, nrow = 2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol = 3, nrow = 2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol = 3, nrow = 2)
p4
```

Now we investigate the influence of prior information on inversion results. This time we focuse on 0.05 noise, 1:3 test. But we change prior information, i.e., geo, we give false information, like mean value too high or low, variance too high or low, range too high or low.

```{r prior_info_influence, cache=TRUE, eval=FALSE}
set.seed(300)
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
qHT1 = blh_synthetic$qHT[1:3]
oHT1 = blh_synthetic$oHT[1:3]
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}
lire = list()
geo_list = list(
  list(me = 2.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0), # low mean
  list(me = 4.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0), # high mean
  list(me = 3.5, var = 0.5, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0), # low var
  list(me = 3.5, var = 2, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0), # high var
  list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 20, nugget = 0), # low range
  list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 80, nugget = 0)  # high range
)
for(i in 1:length(geo_list)){
  geo = geo_list[[i]]
  result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT_noised, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}
### save the result
saveRDS(lire, file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_05_prior-info-test1-3.rds")
```


now plot the prior information influence results.

```{r plot_prior_info_influence, eval=FALSE}
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_05_prior-info-test1-3.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = c("(a) 低均值先验信息",
        "(b) 高均值先验信息",
        "(c) 低方差先验信息",
        "(d) 高方差先验信息",
        "(e) 低相关长度先验信息",
        "(f) 高相关长度先验信息")
for(i in 1:6){
  niterm = length(lire[[i]])
  oHT1 = blh_synthetic$oHT[1:3]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m",
                    p4title = i1[i],
                    p4xlab = "参考lnT",
                    p4ylab = "估计lnT")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol = 3, nrow = 2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol = 3, nrow = 2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol = 3, nrow = 2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol = 3, nrow = 2)
p4
```


It seems there is little influence of prior information on inversion results when 3 tests and 5% noise are given. Now we use only one test to see how prior information affects inversion results.

```{r prior_info_influence_one_test, cache=TRUE, eval=FALSE}
set.seed(300)
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
qHT1 = blh_synthetic$qHT[1]
oHT1 = blh_synthetic$oHT[1]
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}
lire = list()
geo_list = list(
  list(me = 2.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0), # low mean
  list(me = 4.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0), # high mean
  list(me = 3.5, var = 0.5, geomod = "Exp", anis = c(90, 1),  range = 40, nugget = 0), # low var
  list(me = 3.5, var = 2, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0), # high var
  list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 20, nugget = 0), # low range
  list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 80, nugget = 0)  # high range
)
for(i in 1:length(geo_list)){
  geo = geo_list[[i]]
  result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT_noised, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}
### save the result
saveRDS(lire, file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_05_prior-info-test1.rds")
```

```{r plot_prior_info_influence_one_test, eval=FALSE}
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_05_prior-info-test1.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = c("(a) 低均值先验信息",
        "(b) 高均值先验信息",
        "(c) 低方差先验信息",
        "(d) 高方差先验信息",
        "(e) 低相关长度先验信息",
        "(f) 高相关长度先验信息")
for(i in 1:6){
  niterm = length(lire[[i]])
  oHT1 = blh_synthetic$oHT[1]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m",
                    p4title = i1[i],
                    p4xlab = "参考lnT",
                    p4ylab = "估计lnT")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol = 3, nrow = 2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol = 3, nrow = 2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol = 3, nrow = 2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol = 3, nrow = 2)
p4
```

Now we give more details on different prior mean. from me =0.5 to me=5.5, step=1.

```{r detailed_prior_mean_influence, cache=TRUE, eval=FALSE}
set.seed(300)
lire = list()
geo_list = list(
  list(me = 0.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0),
  list(me = 1.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0),
  list(me = 2.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0),
  list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0),
  list(me = 4.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0),
  list(me = 5.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0)
)
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
qHT1 = blh_synthetic$qHT[1:1]
oHT1 = blh_synthetic$oHT[1:1]
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}
for(i in 1:length(geo_list)){
  geo = geo_list[[i]]
  result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT_noised, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}
### save the result
saveRDS(lire, file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_05_prior-info-test1-3-detailed-mean.rds")
```

```{r plot_detailed_prior_mean_influence, eval=FALSE}
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_with-obs-error-sd0_05_prior-info-test1-3-detailed-mean.rds")
i1 = c("(a) 先验均值=0.5",
        "(b) 先验均值=1.5",
        "(c) 先验均值=2.5",
        "(d) 先验均值=3.5",
        "(e) 先验均值=4.5",
        "(f) 先验均值=5.5")
for(i in 1:6){
  niterm = length(lire[[i]])
  oHT1 = blh_synthetic$oHT[1]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m",
                    p4title = i1[i],
                    p4xlab = "参考lnT",
                    p4ylab = "估计lnT")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol = 3, nrow = 2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol = 3, nrow = 2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol = 3, nrow = 2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol = 3, nrow = 2)
p4
```

Now we use the 1:4 true data set  to do the inversion. Using default prior, and then we use the inverted T to do the prediction on test 5-6.

```{r prediction_using_inverted_T, cache=TRUE, eval=FALSE}
set.seed(300)
data('oHT')
data('qHT')
# get 1-4 test data
oHT1 = oHT[c(1,3,5,7)]
qHT1 = qHT[c(1,3,5,7)]

geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0)
result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT1, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
# save the result.
saveRDS(result, file = "E:/HydroTomo/result_real_data_test1-4.rds")
```

```{r plot_real_data_inversion_results, eval=FALSE}
### inverseplot, without trueK.
result = readRDS(file = "E:/HydroTomo/result_real_data_test1-4.rds")
data('oHT')
oHT1 = oHT[c(1,3,5,7)]
p= inversePlot(niterm=length(result),grid=grid,iterdf = result,oHT = oHT1,trueK=NULL,
                    p1title = "(a) 1-4次抽水试验反演结果",
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = "(b) 1-4次抽水试验反演结果",
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = "(c) 1-4次抽水试验反演结果",
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m")
p$lnk
p$varlnk
p$headscatter
```

```{r plot_real_data_prediction_results, eval=FALSE}
### now we use that to do the prediction on test 5-6
niterm = length(result)
invertedK = result[[niterm]]$meanT
TT_inv = exp(invertedK)
# test 5-6
qHT_pred = qHT[c(9,11)]
oHT_pred = oHT[c(9,11)]
pred_results = list()
for(i in 1:length(qHT_pred)){
  Qinf = qHT_pred[[i]]
  simh <- Fsteady2dsim(TT=TT_inv,Qinf=Qinf,grid=grid,lrw=320000)$solution
  oinf <- oHT_pred[[i]]
  pred_df = data.frame(observed = oinf$data,
                       simulated = samData(grid=grid,Oinf=oinf,h=simh)$data)
  pred_results[[i]] = pred_df
}

# make pred_results a long data frame
library(dplyr)
pred_results_df = dplyr::bind_rows(pred_results,.id='test_id')
## plot the head scatter
p_pred = ggplot(pred_results_df, aes(x=-observed, y=-simulated)) +
  geom_point(color='blue', alpha=0.5) +
  geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
  labs(title='预测结果散点图（使用1-4次试验反演的T）', x='观测降深/m', y='模拟降深/m') +
  coord_fixed() +
  theme_minimal()
p_pred

### save the result
saveRDS(pred_results_df, file = "E:/HydroTomo/prediction_using_inverted_T_test5-6.rds")
```


Now we check using 1-4 test inverted T to predict test 5-6 results.

```{r inverse_using_1_to_4_prediction_results, eval=FALSE}
set.seed(300)
data('oHT')
data('qHT')
# get 1-4 test data
oHT1 = oHT[c(1,3,5,7)]
qHT1 = qHT[c(1,3,5,7)]
geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0)
lire = list()
for(i in 1:4){
  qHT_pred = qHT1[1:i]
  oHT_pred = oHT1[1:i]
  result <- Finverse(grid =grid, qHT = qHT_pred, oHT = oHT_pred, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}

### save the result
saveRDS(lire, file = "E:/HydroTomo/result_real_data_test1-4_incremental_test.rds")
```

now we also plot the incremental test results.
```{r plot_real_data_incremental_inversion_results, eval=FALSE}
data('oHT')
data('qHT')
lire = readRDS(file = "E:/HydroTomo/result_real_data_test1-4_incremental_test.rds")
p1k = list()
p2k = list()
p3k = list()
i1 = paste0("(",letters[1:4],") " ,1:4,"次抽水试验")
for(i in 1:4){
  niterm = length(lire[[i]])
  oHT1 = oHT[c(1,3,5,7)][1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=NULL,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol=2, nrow=2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol=2, nrow=2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol=2, nrow=2)
p3
```

Now we use the 1-4 test inverted T to predict test 5-6 results.

```{r plot_real_data_incremental_prediction_results, eval=FALSE}
### now we use that to do the prediction on test 5-6

plist = list()
# test 5-6
qHT_pred = qHT[c(9,11)]
oHT_pred = oHT[c(9,11)]
for(j in 1:4){
  pred_results = list()
  niterm = length(lire[[j]])
  invertedK = lire[[j]][[niterm]]$meanT
  TT_inv = exp(invertedK)
for(i in 1:length(qHT_pred)){
  Qinf = qHT_pred[[i]]
  simh <- Fsteady2dsim(TT=TT_inv,Qinf=Qinf,grid=grid,lrw=320000)$solution
  oinf <- oHT_pred[[i]]
  pred_df = data.frame(observed = oinf$data,
                       simulated = samData(grid=grid,Oinf=oinf,h=simh)$data)
  pred_results[[i]] = pred_df
}
# make pred_results a long data frame
library(dplyr)
pred_results_df = dplyr::bind_rows(pred_results,.id='test_id')
## plot the head scatter
p_pred = ggplot(pred_results_df, aes(x=-observed, y=-simulated)) +
  geom_point(color='blue', alpha=0.5) +
  geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
  labs(title='预测结果散点图（使用1-4次试验反演的T，增量式）', x='观测降深/m', y='模拟降深/m') +
  theme_minimal()
p_pred

plist[[j]] = pred_results_df
}

### save the result
saveRDS(plist, file = "E:/HydroTomo/prediction_using_inverted_T_test5-6_incremental.rds")
```

```{r plot_incremental_prediction_results, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_test5-6_incremental.rds")
p_pred = list()
for(i in 1:4){
  pred_results_df = plist[[j]]
  p_pred[[i]] = ggplot(pred_results_df, aes(x=-observed, y=-simulated)) +
    geom_point(color='blue', alpha=0.5) +
    geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
    labs(title=paste0("(",letters[j],") 使用1-",i,"次试验反演"), x='观测降深/m', y='模拟降深/m') +
    theme_minimal()
}
p_all = ggpubr::ggarrange(plotlist = p_pred, ncol=2, nrow=2)
p_all
```


Now we  using 1-4 test in blh_syn_best_truth to inverted T .

```{r invert_blh_syn_best_truth, cache=TRUE, eval=FALSE}
set.seed(300)
data('blh_syn_best_truth')
# it seems the blh_syn_best_truth is from blh_synthetic.rmd, set.seed(200)
grid = blh_syn_best_truth$grid
trueK = blh_syn_best_truth$trueK
qHT1 = blh_syn_best_truth$qHT
oHT1 = blh_syn_best_truth$oHT
# add 0.05 sd noise
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}

geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0)


lire = list()
for(i in 1:4){
  qHT_pred = qHT1[1:i]
  oHT_pred = oHT_noised[1:i]
  result <- Finverse(grid =grid, qHT = qHT_pred, oHT = oHT_pred, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}
# save the result.
saveRDS(lire, file = "E:/HydroTomo/result_blh_syn_best_truth_test1-4.rds")
```


```{r plot_result_blh_syn_best_truth, cache=TRUE, eval=FALSE}
set.seed(300)
# add 0.05 sd noise
data('blh_syn_best_truth')
grid = blh_syn_best_truth$grid
trueK = blh_syn_best_truth$trueK
qHT1 = blh_syn_best_truth$qHT
oHT1 = blh_syn_best_truth$oHT
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}

lire = readRDS(file = "E:/HydroTomo/result_blh_syn_best_truth_test1-4.rds")
p1k = list()
p2k = list()
p3k = list()
i1 = paste0("(",letters[1:4],") " ,1:4,"次抽水试验，观测数据含5%噪声")
for(i in 1:4){
  niterm = length(lire[[i]])
  oHT_pred = oHT_noised[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT_pred,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol=2, nrow=2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol=2, nrow=2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol=2, nrow=2)
p3
```


Now we use the 1-4 test inverted T to predict test 5-6 results and plot.

```{r plot_blh_syn_best_truth_prediction_results, eval=FALSE}
### now we use that to do the prediction on test 5-6
plist = list()
# test 5-6
qHT_pred = qHT1[c(5,6)]
oHT_pred = oHT_noised[c(5,6)]
for(j in 1:4){
  pred_results = list()
  niterm = length(lire[[j]])
  invertedK = lire[[j]][[niterm]]$meanT
  TT_inv = exp(invertedK)
  for(i in 1:length(qHT_pred)){
    Qinf = qHT_pred[[i]]
    simh <- Fsteady2dsim(TT=TT_inv,Qinf=Qinf,grid=grid,lrw=320000)$solution
    oinf <- oHT_pred[[i]]
    pred_df = data.frame(observed = oinf$data,
                         simulated = samData(grid=grid,Oinf=oinf,h=simh)$data)
    pred_results[[i]] = pred_df
  }
  # make pred_results a long data frame
  library(dplyr)
  pred_results_df = dplyr::bind_rows(pred_results,.id='test_id')
  ## plot the head scatter
  p_pred = ggplot(pred_results_df, aes(x=observed, y=simulated)) +
    geom_point(color='blue', alpha=0.5) +
    geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
    labs(title='预测结果散点图（使用1-4次试验反演的T）', x='观测降深/m', y='模拟降深/m') +
    theme_minimal()
  p_pred
  
  plist[[j]] = pred_results_df
}
### save the result
saveRDS(plist, file = "E:/HydroTomo/prediction_using_inverted_T_blh_syn_best_truth_test5-6.rds")
```

```{r plot_blh_syn_best_truth_prediction_results, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_blh_syn_best_truth_test5-6.rds")
p_pred = list()
for(i in 1:4){
  pred_results_df = plist[[j]]
  p_pred[[i]] = ggplot(pred_results_df, aes(x=observed, y=simulated)) +
    geom_point(color='blue', alpha=0.5) +
    geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
    labs(title=paste0("(",letters[j],") 使用1-",j,"次i验反演的T预测5-6次试验结果"), x='观测降深/m', y='模拟降深/m') +
    theme_minimal()
}
p_all = ggpubr::ggarrange(plotlist = p_pred, ncol=2, nrow=2)
p_all
```


# generate a random field using range =20

```{r generate_random_field_range20, cache=TRUE, eval=FALSE}
set.seed(400)
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
qHT1 = blh_synthetic$qHT
oHT1 = blh_synthetic$oHT
geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 20, nugget = 0)
trueK <- random2d(nsim=1,grid=grid,geo = geo)
Plotparameter2d(TT=trueK,iflog=T)
```

Now we use this T field to generate synethetic data for test 1-6.

```{r generate_synthetic_data_range20, cache=TRUE, eval=FALSE}
set.seed(400)
data('blh_synthetic')
grid = blh_synthetic$grid
qHT1 = blh_synthetic$qHT
oHT1 = blh_synthetic$oHT
TT <- trueK$Tp
for (i in 1:length(oHT1)){
  Qinf <- qHT1[[i]]
  trueh <- Fsteady2dsim(TT=TT,Qinf=Qinf,grid=grid,lrw=320000)$solution
  oinf <- oHT1[[i]]
  oinf = samData(grid = grid,Oinf = oinf,h=trueh)
  oHT1[[i]] <- oinf[,c('data','x','y')]
}

dfoHT1 = do.call(rbind, oHT1)
mean_oHT1 = mean(dfoHT1$data)
print(paste("The mean value of oHT1 is ", round(mean_oHT1,4)))

```
We save the data as 'blh_synthetic_range20' object.

```{r save_blh_synthetic_range20, cache=TRUE, eval=FALSE}
blh_synthetic_range20 = list()
blh_synthetic_range20$grid = grid
blh_synthetic_range20$trueK = TT
blh_synthetic_range20$qHT = qHT1
blh_synthetic_range20$oHT = oHT1
saveRDS(blh_synthetic_range20, file = "E:/HydroTomo/blh_synthetic_range20.rds")
```

we use the synthetic data (add 0.1 sd noise ) to do the inversion.

```{r invert_synthetic_data_range20, cache=TRUE, eval=FALSE}
set.seed(400)
blh_synthetic_range20 = readRDS(file = "E:/HydroTomo/blh_synthetic_range20.rds")
grid = blh_synthetic_range20$grid
trueK = blh_synthetic_range20$trueK
qHT1 = blh_synthetic_range20$qHT
oHT1 = blh_synthetic_range20$oHT
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.1*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}

geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 20, nugget = 0)

# do incremental inversion using data test 1, test 1-2, test 1-3, ..., test 1-4.
lire = list()
for(i in 1:4){
  qHT_pred = qHT1[1:i]
  oHT_pred = oHT_noised[1:i]
  result <- Finverse(grid =grid, qHT = qHT_pred, oHT = oHT_pred, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}

# save the result.
saveRDS(lire, file = "E:/HydroTomo/result_blh_synthetic_range20_with-obs-error-sd0_1.rds")
```


plot the inverse result.

```{r plot_inversion_results_range20, eval=FALSE}
blh_synthetic_range20 = readRDS(file = "E:/HydroTomo/blh_synthetic_range20.rds")
grid = blh_synthetic_range20$grid
trueK = blh_synthetic_range20$trueK
trueK = data.frame(x=grid$grid$x, y=grid$grid$y, Tp=trueK)
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_range20_with-obs-error-sd0_1.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = paste0("(",letters[1:4],") " ,1:4,"次抽水试验，观测数据含10%噪声")
for(i in 1:4){
  niterm = length(lire[[i]])
  oHT_pred = blh_synthetic_range20$oHT[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT_pred,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol=2, nrow=2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol=2, nrow=2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol=2, nrow=2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol=2, nrow=2)
p4
```


Now we use the 1-4 test inverted T to predict test 5-6 results and plot.

```{r plot_prediction_results_range20, eval=FALSE}
### now we use that to do the prediction on test 5-6
plist = list()
# test 5-6
qHT_pred = blh_synthetic_range20$qHT[c(5,6)]
oHT_pred = blh_synthetic_range20$oHT[c(5,6)]
for(j in 1:4){
  pred_results = list()
  niterm = length(lire[[j]])
  invertedK = lire[[j]][[niterm]]$meanT
  TT_inv = exp(invertedK)
  for(i in 1:length(qHT_pred)){
    Qinf = qHT_pred[[i]]
    simh <- Fsteady2dsim(TT=TT_inv,Qinf=Qinf,grid=grid,lrw=320000)$solution
    oinf <- oHT_pred[[i]]
    pred_df = data.frame(observed = oinf$data,
                         simulated = samData(grid=grid,Oinf=oinf,h=simh)$data)
    pred_results[[i]] = pred_df
  }
  # make pred_results a long data frame
  library(dplyr)
  pred_results_df = dplyr::bind_rows(pred_results,.id='test_id')
  ## plot the head scatter
  p_pred = ggplot(pred_results_df, aes(x=observed, y=simulated)) +
    geom_point(color='blue', alpha=0.5) +
    geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
    labs(title='预测结果散点图（使用1-4次试验反演的T）', x='观测降深/m', y='模拟降深/m') +
    theme_minimal()
  p_pred
  plist[[j]] = pred_results_df
}
### save the result
saveRDS(plist, file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6.rds")
```

```{r plot_prediction_results_range20, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6.rds")
p_pred = list()
for(i in 1:4){
  pred_results_df = plist[[j]]
  p_pred[[i]] = ggplot(pred_results_df, aes(x=observed, y=simulated)) +
    geom_point(color='blue', alpha=0.5) +
    geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
    labs(title=paste0("(",letters[j],") 使用1-",j,"次i验反演的T预测5-6次试验结果"), x='观测降深/m', y='模拟降深/m') +
    theme_minimal()
}
p_all = ggpubr::ggarrange(plotlist = p_pred, ncol=2, nrow=2)
p_all
```
Now we calculate the rmse for the four cases.

```{r calculate_rmse_range20, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6.rds")
mc = data.frame(test_id = character(),
                rmse = numeric(),
                r2 = numeric())
for(i in 1:4){
  pred_results_df = plist[[i]]
  rmse = sqrt(mean((pred_results_df$observed - pred_results_df$simulated)^2))
  r2 = cor(pred_results_df$observed, pred_results_df$simulated)^2
  mc = rbind(mc, data.frame(test_id = paste0("1-",i,"次试验"), rmse = rmse, r2 = r2))
}
mc
```


Now we reduce the number of observation in each test.

```{r invert_synthetic_data_range20_fewer_obs, cache=TRUE, eval=FALSE}
set.seed(400)
blh_synthetic_range20 = readRDS(file = "E:/HydroTomo/blh_synthetic_range20.rds")
grid = blh_synthetic_range20$grid
trueK = blh_synthetic_range20$trueK
qHT1 = blh_synthetic_range20$qHT
oHT1 = blh_synthetic_range20$oHT
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.1*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}

oHT_reduced = list()
for(i in 1:length(oHT_noised)){
  oinf = oHT_noised[[i]]
  # sample 40% of the data
  set.seed(400 + i)
  sampled_indices = sample(1:nrow(oinf), size = floor(0.4 * nrow(oinf)), replace = FALSE)
  oinf_reduced = oinf[sampled_indices, ]
  oHT_reduced[[i]] = oinf_reduced
}

geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 20, nugget = 0)
# do incremental inversion using data test 1, test 1-2, test 1-3, ..., test 1-4.
lire = list()
for(i in 1:4){
  qHT_pred = qHT1[1:i]
  oHT_pred = oHT_reduced[1:i]
  result <- Finverse(grid =grid, qHT = qHT_pred, oHT = oHT_pred, nsim=50, itermax=10, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}
# save the result.
saveRDS(lire, file = "E:/HydroTomo/result_blh_synthetic_range20_with-obs-error-sd0_1_fewer-obs.rds")
```

```{r plot_inversion_results_range20_fewer_obs, eval=FALSE}
blh_synthetic_range20 = readRDS(file = "E:/HydroTomo/blh_synthetic_range20.rds")
grid = blh_synthetic_range20$grid
trueK = blh_synthetic_range20$trueK
trueK = data.frame(x=grid$grid$x, y=grid$grid$y, Tp=trueK)
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_range20_with-obs-error-sd0_1_fewer-obs.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = paste0("(",letters[1:4],") " ,1:4,"次抽水试验，观测数据含10%噪声，观测点数40%")
for(i in 1:4){
  niterm = length(lire[[i]])
  oHT_pred = oHT_reduced[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT_pred,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol=2, nrow=2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol=2, nrow=2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol=2, nrow=2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol=2, nrow=2)
p4
```

Now we use the 1-4 test inverted T to predict test 5-6 results and plot.
```{r plot_prediction_results_range20_fewer_obs, eval=FALSE}
### now we use that to do the prediction on test 5-6
plist = list()
# test 5-6
qHT_pred = blh_synthetic_range20$qHT[c(5,6)]
oHT_pred = blh_synthetic_range20$oHT[c(5,6)]
for(j in 1:4){
  pred_results = list()
  niterm = length(lire[[j]])
  invertedK = lire[[j]][[niterm]]$meanT
  TT_inv = exp(invertedK)
  for(i in 1:length(qHT_pred)){
    Qinf = qHT_pred[[i]]
    simh <- Fsteady2dsim(TT=TT_inv,Qinf=Qinf,grid=grid,lrw=320000)$solution
    oinf <- oHT_pred[[i]]
    pred_df = data.frame(observed = oinf$data,
                         simulated = samData(grid=grid,Oinf=oinf,h=simh)$data)
    pred_results[[i]] = pred_df
  }
  # make pred_results a long data frame
  library(dplyr)
  pred_results_df = dplyr::bind_rows(pred_results,.id='test_id')

  plist[[j]] = pred_results_df
}
### save the result
saveRDS(plist, file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6_fewer-obs.rds")
```


```{r plot_prediction_results_range20_fewer_obs, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6_fewer-obs.rds")
p_pred = list()
for(i in 1:4){
  pred_results_df = plist[[i]]
  p_pred[[i]] = ggplot(pred_results_df, aes(x=observed, y=simulated)) +
    geom_point(color='blue', alpha=0.5) +
    geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
    labs(title=paste0("(",letters[i],") 使用1-",i,"次i验反演的T预测5-6次试验结果，观测点数减半"), x='观测降深/m', y='模拟降深/m') +
    theme_minimal()
}
p_all = ggpubr::ggarrange(plotlist = p_pred, ncol=2, nrow=2)
p_all
```
Now calculate rmse and r2.

```{r calculate_rmse_range20_less_obs, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6_fewer-obs.rds")
mc = data.frame(test_id = character(),
                rmse = numeric(),
                r2 = numeric())
for(i in 1:4){
  pred_results_df = plist[[i]]
  rmse = sqrt(mean((pred_results_df$observed - pred_results_df$simulated)^2))
  r2 = cor(pred_results_df$observed, pred_results_df$simulated)^2
  mc = rbind(mc, data.frame(test_id = paste0("1-",i,"次试验"), rmse = rmse, r2 = r2))
}
mc
```



Now we reduce the number of observation and set noise = 0 in each test.

```{r invert_synthetic_data_range20_fewer_obs, cache=TRUE, eval=FALSE}
set.seed(400)
blh_synthetic_range20 = readRDS(file = "E:/HydroTomo/blh_synthetic_range20.rds")
grid = blh_synthetic_range20$grid
trueK = blh_synthetic_range20$trueK
qHT1 = blh_synthetic_range20$qHT
oHT1 = blh_synthetic_range20$oHT
oHT_noised <- oHT1
for (i in 1:length(oHT_noised)){
  noise = rnorm(n=nrow(oHT_noised[[i]]), mean=0, sd=0.05*sd(oHT_noised[[i]]$data))
  oHT_noised[[i]]$data <- oHT_noised[[i]]$data + noise
}

oHT_reduced = list()
for(i in 1:length(oHT_noised)){
  oinf = oHT_noised[[i]]
  # sample 40% of the data
  set.seed(400 + i)
  sampled_indices = sample(1:nrow(oinf), size = floor(0.3 * nrow(oinf)), replace = FALSE)
  oinf_reduced = oinf[sampled_indices, ]
  oHT_reduced[[i]] = oinf_reduced
}

geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 20, nugget = 0)
# do incremental inversion using data test 1, test 1-2, test 1-3, ..., test 1-4.
lire = list()
for(i in 1:4){
  qHT_pred = qHT1[1:i]
  oHT_pred = oHT_reduced[1:i]
  result <- Finverse(grid =grid, qHT = qHT_pred, oHT = oHT_pred, nsim=50, itermax=15, lrw=320000,
                     geo = geo)
  lire[[i]] = result
}
# save the result.
saveRDS(lire, file = "E:/HydroTomo/result_blh_synthetic_range20_with-obs-error-sd0_0_fewerer-obs.rds")
```

```{r plot_inversion_results_range20_fewer_obs_noise0.0, eval=FALSE}
blh_synthetic_range20 = readRDS(file = "E:/HydroTomo/blh_synthetic_range20.rds")
grid = blh_synthetic_range20$grid
trueK = blh_synthetic_range20$trueK
trueK = data.frame(x=grid$grid$x, y=grid$grid$y, Tp=trueK)
lire = readRDS(file = "E:/HydroTomo/result_blh_synthetic_range20_with-obs-error-sd0_0_fewerer-obs.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = paste0("(",letters[1:4],") " ,1:4,"次抽水试验，观测数据含0%噪声，观测点数40%")
for(i in 1:4){
  niterm = length(lire[[i]])
  oHT_pred = oHT_reduced[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT_pred,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol=2, nrow=2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol=2, nrow=2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol=2, nrow=2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol=2, nrow=2)
p4
```

Now we use the 1-4 test inverted T to predict test 5-6 results and plot.
```{r plot_prediction_results_range20_fewer_obs_noise_0.0, eval=FALSE}
### now we use that to do the prediction on test 5-6
plist = list()
# test 5-6
qHT_pred = blh_synthetic_range20$qHT[c(5,6)]
oHT_pred = blh_synthetic_range20$oHT[c(5,6)]
for(j in 1:4){
  pred_results = list()
  niterm = length(lire[[j]])
  invertedK = lire[[j]][[niterm]]$meanT
  TT_inv = exp(invertedK)
  for(i in 1:length(qHT_pred)){
    Qinf = qHT_pred[[i]]
    simh <- Fsteady2dsim(TT=TT_inv,Qinf=Qinf,grid=grid,lrw=320000)$solution
    oinf <- oHT_pred[[i]]
    pred_df = data.frame(observed = oinf$data,
                         simulated = samData(grid=grid,Oinf=oinf,h=simh)$data)
    pred_results[[i]] = pred_df
  }
  # make pred_results a long data frame
  library(dplyr)
  pred_results_df = dplyr::bind_rows(pred_results,.id='test_id')

  plist[[j]] = pred_results_df
}
### save the result
saveRDS(plist, file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6_fewerer-obs_noise_0.0.rds")
```


```{r plot_prediction_results_range20_fewer_obs_noise_0.0, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6_fewerer-obs_noise_0.0.rds")
p_pred = list()
for(i in 1:4){
  pred_results_df = plist[[i]]
  p_pred[[i]] = ggplot(pred_results_df, aes(x=observed, y=simulated)) +
    geom_point(color='blue', alpha=0.5) +
    geom_abline(slope=1, intercept=0, color='red', linetype='dashed') +
    labs(title=paste0("(",letters[i],") 使用1-",i,"次i验反演的T预测5-6次试验结果，观测点数减半"), x='观测降深/m', y='模拟降深/m') +
    theme_minimal()
}
p_all = ggpubr::ggarrange(plotlist = p_pred, ncol=2, nrow=2)
p_all
```
Now calculate rmse and r2.

```{r calculate_rmse_range20_less_obs_noise_0.0, eval=FALSE}
plist = readRDS(file = "E:/HydroTomo/prediction_using_inverted_T_blh_synthetic_range20_test5-6_fewerer-obs_noise_0.0.rds")
mc = data.frame(test_id = character(),
                rmse = numeric(),
                r2 = numeric())
for(i in 1:4){
  pred_results_df = plist[[i]]
  rmse = sqrt(mean((pred_results_df$observed - pred_results_df$simulated)^2))
  r2 = cor(pred_results_df$observed, pred_results_df$simulated)^2
  mc = rbind(mc, data.frame(test_id = paste0("1-",i,"次试验"), rmse = rmse, r2 = r2))
}
mc
```
