---
title: "hydraulic tomography incremental paper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{hydraulic tomography incremental paper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is to create the paper data and figures.

```{r setup}
library(HydroTomo)
```

# Fig. 1

This figure draw the synthetic case of hydraulic tomography. In figure a, we generate a random field, and gives the 6 pumping well locations. In figure b, we show the simulated drawdown data for test A-1 and also the location of observation wells in test A-1.

```{r figure1}
set.seed(2025)
data('qHT')
library(dplyr)
qHTdf = dplyr::bind_rows(qHT,.id='id')
qHTdf = qHTdf %>% dplyr::filter(id %in% c('A1-1','A2-1','A3-1','A4-1','A5-1','A6-1'))

domain=c(80,80,0,80,0,80)
grid <- GenGrid(domain=domain)
# exp(3.5)=33.12 m2/d means very large T values in sand and gravel aquifer.
geo = list(me = 3.2, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0)
trueK <- random2d(nsim=1,grid=grid,geo = geo)
trueK$Tp = log(trueK$Tp)
# make sure chinese font show correctly
library(ggplot2)
library(showtext)
showtext_auto()
  p1 = ggplot() +
    geom_tile(data = trueK, aes(x, y, fill = Tp)) +
    scale_fill_viridis_c(option = "D")+
    geom_point(data = qHTdf, aes(x = x, y = y), color = "white", size = 3, shape=17)+
   geom_text(data = qHTdf, aes(x = x, y = y, label = id), vjust = -1, color = "black")+
    labs(title = "(a) 参考lnT及抽水井", x = "X/m", y = "Y/m", fill = "log(T)")+
    theme_minimal()
  p1
```

```{r figure1b}
# simulate drawdown data for test A1-1
TT <- exp(trueK$Tp)
Qinf = qHT[['A1-1']]
data('oHT')
library(dplyr)
oHTdf = dplyr::bind_rows(oHT,.id='id')
oHTdf = oHTdf %>% dplyr::filter(id %in% c('A1-1'))
s = Fsteady2dsim(TT=TT,Qinf=Qinf,grid=grid,lrw=320000)

  p2 = ggplot() +
    geom_tile(data = s, aes(x, y, fill = -solution)) +
    scale_fill_viridis_c(option = "D")+
    geom_point(data = oHTdf, aes(x = x, y = y), color = "red",fill="white", size = 3, shape=21)+
    labs(title = "(b) 模拟降深及观测井", x = "X/m", y = "Y/m", fill = "降深/m")+
    theme_minimal()
  p2
```  
Now we use ggarrange to make the p1 p2 together.

```{r figure1c, fig.width=10, fig.height=5}
library(ggpubr)

p = ggarrange(p1, p2, ncol = 2, nrow = 1)

p
```

# Generating synthetic data

We then simulate steady-state drawdown data for a pumping test using the true transmissivity field.

```{r generate_synthetic_data, eval=FALSE}
# TT <- 
qHT1 = qHT[c(1,3,5,7,9,11)]
oHT1 = oHT[c(1,3,5,7,9,11)]
### loop over the list to update the value in oHT1.

for (i in 1:length(oHT1)){
  Qinf <- qHT1[[i]]
  trueh <- Fsteady2dsim(TT=TT,Qinf=Qinf,grid=grid,lrw=320000)$solution
  oinf <- oHT1[[i]]
  oinf = samData(grid = grid,Oinf = oinf,h=trueh)
  oHT1[[i]] <- oinf[,c('data','x','y')]
}

dfoHT1 = do.call(rbind,oHT1)
mean_oHT1 = mean(dfoHT1$data)
print(paste("The mean value of oHT1 is ", round(mean_oHT1,4)))
```

save the best synthetic data.

```{r save_data, eval=FALSE}
blh_syn_best_truth <- list(trueK=trueK, qHT=qHT1, oHT=oHT1, grid=grid)
usethis::use_data(blh_syn_best_truth, overwrite = TRUE)
```
Now we use "blh_synthetic" data to do a sequence of inversion and save the result.

```{r perform_inversion, cache=TRUE, eval=FALSE}
data('blh_synthetic')
grid = blh_synthetic$grid
lire = list()
# we select qHT1 and oHT1 from blh_synthetic, first test1, then 1,2, then 1,2,3
for(i in 1:6){
  qHT1 = blh_synthetic$qHT[1:i]
  oHT1 = blh_synthetic$oHT[1:i]
  result <- Finverse(grid =grid, qHT = qHT1, oHT = oHT1, nsim=50, itermax=10, lrw=320000,
                     geo = list(me = 3.5, var = 1, geomod = "Exp", anis = c(90, 1), range = 40, nugget = 0))
  lire[[i]] = result
}
### save the result
saveRDS(lire, file = "lire_blh_synthetic_no-obs-error.rds")
```

# Plotting inversion results

```{r plot_inversion_results, eval=FALSE}
data('blh_synthetic')
grid = blh_synthetic$grid
trueK = blh_synthetic$trueK
lire = readRDS(file = "E:/HydroTomo/lire_blh_synthetic_no-obs-error.rds")
p1k = list()
p2k = list()
p3k = list()
p4k = list()
i1 = paste0("(",letters[1:6],") " ,1:6,"次抽水试验")
for(i in 1:6){
  niterm = length(lire[[i]])
  oHT1 = blh_synthetic$oHT[1:i]
  tmp = inversePlot(niterm=niterm,grid=grid,iterdf = lire[[i]],oHT = oHT1,trueK=trueK,
                    p1title = i1[i],
                    p1z = "ln(T) [L2/T]",
                    p1xlab = "X/m",
                    p1ylab = "Y/m",
                    p2title = i1[i],
                    p2z = "Var[ln(T)] [L2/T]",
                    p2xlab = "X/m",
                    p2ylab = "Y/m",
                    p3title = i1[i],
                    p3xlab = "模拟降深/m",
                    p3ylab = "观测降深/m",
                    p4title = i1[i],
                    p4xlab = "参考lnT",
                    p4ylab = "估计lnT")
  p1k[[i]] = tmp$lnk
  p2k[[i]] = tmp$varlnk
  p3k[[i]] = tmp$headscatter
  p4k[[i]] = tmp$lnkscatter
}
p1 = ggpubr::ggarrange(plotlist = p1k, ncol = 3, nrow = 2)
p1
p2 = ggpubr::ggarrange(plotlist = p2k, ncol = 3, nrow = 2)
p2
p3 = ggpubr::ggarrange(plotlist = p3k, ncol = 3, nrow = 2)
p3
p4 = ggpubr::ggarrange(plotlist = p4k, ncol = 3, nrow = 2)
p4
```
